### 2022-02-27 outputs/paper/paper.Rmd
add more data

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Lantao Cui
  - Huakun Shen
  - Sijian Wu
thanks: "Code and data are available at: https://github.com/cuilantao/STA304-Paper-2"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Between 1980 and the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44. However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years. In this paper, we aim to find out the reason behind the decline, all of the data are collected from the CDC Vital Statistics Births Reports. Our findings can help to understand the reason behind the drop. "
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(here)
library(ggplot2)
library(ggpubr)
```

# Introduction

The Great Recession has been a world-wide crisis that reconstructed the economy, and was deep, synchronous and global. Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent and that of Japan by more than 8 percent. The impact on international trade has been severe that all G20 countries experienced serious reductions in trade[@recession]. Although it is difficult to compare across countries since policy context, society norms and cultural preferences tend to differ, it is never possible to separately consider the effects on each country even after recovery.

In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas. Another impact from depressed economic condition is delayed childbearing, and others to hasten it. The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility]. The significant consequence on fertility rate would make us wonder how severe is America being affected by the great recession. However,  by comparing currency devaluations, equity market declines, and rising sovereign bond spreads, the U.S. is the third leas affected country[@rank].

Prior research suggests that stress may be linked to riskier sextual activity. Financial hardship and poverty can negatively affect cognitive function, and shorten the time horizons over which individuals make cost-benefit determinations. Alternatively, fertility may be unaffected by economic conditions if cultural norms related to the context and timing of births are particularly influential. For example, non-marital births are more socially consequential for some racial and ethnic groups, and in some geographic areas. For some individuals, these social norms may outweigh any economic considerations[@fertility]. The same consideration of religious belief could also be applied to other countries.

# Data

We started our analysis by using R [@r], dplyr[@dplyr], tidyverse[@tidyverse]. Graphs are generated using ggpubr[@ggpubr] and ggplot2[@ggplot2].
```{r, include=FALSE}
#importing the Birth Rate 1950-2022 of the least affected countries: China, Japan, and the United States from https://www.macrotrends.net/countries/USA/united-states/birth-rate
#Then modify the 'date' variable of each into Date format for graphs by using as.Date
china_data <- read_csv(here::here("inputs/data/china-population-2022-02-25.csv"))
china_data$date <- as.Date(china_data$date)
str(china_data)

japan_data <- read.csv(here::here("inputs/data/japan-population-2022-02-25.csv"))
japan_data$date <- as.Date(japan_data$date)
str(japan_data)

us_data <- read_csv(here::here("inputs/data/united-states-population-2022-02-26.csv"))
us_data$date <- as.Date(us_data$date)
str(us_data)

sa_data <- read_csv(here::here("inputs/data/south-africa-population-2022-02-26.csv"))
sa_data$date <- as.Date(sa_data$date)
str(sa_data)
```

```{r,echo=FALSE}
#Generating time series graphs of births per 1000 people for each least affected country

china <- ggplot(china_data,aes(x=date,y= `Births per 1000 People`)) + 
   labs(title = "China",
        x = "Year",
        y = "Births per 1000 People ") +
  theme_minimal() +
  geom_line()

japan <- ggplot(japan_data,aes(x=date,y= Births.per.1000.People)) + 
   labs(title = "Japan",
        x = "Year",
        y = "Births per 1000 People ") +
  theme_minimal() +
  geom_line()

us <- ggplot(us_data,aes(x=date,y= `Births per 1000 People`)) + 
   labs(title = "The United States",
        x = "Year",
        y = "Births per 1000 People ") +
  theme_minimal() +
  geom_line()

sa <- ggplot(sa_data,aes(x=date,y= `Births per 1000 People`)) + 
   labs(title = "South Africa",
        x = "Year",
        y = "Births per 1000 People ") +
  theme_minimal() +
  geom_line()

#Arranging them on the same page
ggarrange(china, japan, us, sa + rremove("x.text"),
          ncol = 2, nrow = 2)
```

```{r, include=FALSE}
#importing the Birth Rate 1950-2022 of the most affected countries: Ukraine, Argentina, Hungary, and Poland from https://www.macrotrends.net/countries/USA/united-states/birth-rate
#Then modify the 'date' variable of each into Date format for graphs by using as.Date
ukraine_data <- read_csv(here::here("inputs/data/ukraine-population-2022-02-26.csv"))
ukraine_data$date <- as.Date(ukraine_data$date)
str(ukraine_data)

argentina_data <- read_csv(here::here("inputs/data/argentina-population-2022-02-26.csv"))
argentina_data$date <- as.Date(argentina_data$date)

hungary_data <- read_csv(here::here("inputs/data/hungary-population-2022-02-27.csv"))
hungary_data$date <- as.Date(hungary_data$date)

poland_data <- read_csv(here::here("inputs/data/poland-population-2022-02-26.csv"))
poland_data$date <- as.Date(poland_data$date)

```

```{r, echo=FALSE}
#Generating time series graphs of births per 1000 people for each most affected country
ukraine <- ggplot(ukraine_data,aes(x=date,y= `Births per 1000 People`)) + 
   labs(title = "Ukraine",
        x = "Year",
        y = "Births per 1000 People ") +
  theme_minimal() +
  geom_line()

argentina <- ggplot(argentina_data,aes(x=date,y= `Births per 1000 People`)) + 
   labs(title = "Argentina",
        x = "Year",
        y = "Births per 1000 People ") +
  theme_minimal() +
  geom_line()

hungary <- ggplot(hungary_data,aes(x=date,y= `Births per 1000 People`)) + 
   labs(title = "Hungary",
        x = "Year",
        y = "Births per 1000 People ") +
  theme_minimal() +
  geom_line()

poland <- ggplot(poland_data,aes(x=date,y= `Births per 1000 People`)) + 
   labs(title = "Poland",
        x = "Year",
        y = "Births per 1000 People ") +
  theme_minimal() +
  geom_line()

ggarrange(ukraine, argentina, hungary, poland + rremove("x.text"),
          ncol = 2, nrow = 2)
```

# Model

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

### 2022-02-27 outputs/paper/paper.Rmd
Add images to rmd

@@
 author: 
-  - Lantao Cui
-  - Huakun Shen
-  - Sijian Wu
-thanks: "Code and data are available at: https://github.com/cuilantao/STA304-Paper-2"
+  - First author
+  - Another author
+thanks: "Code and data are available at: LINK."
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "Between 1980 and the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
-However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years.
-In this paper, we aim to find out the reason behind the decline, all of the data are collected from the CDC Vital Statistics Births Reports.
-Our findings can help to understand the reason behind the drop.
-"
+abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
 output:
@@
 bibliography: references.bib
+header-includes:
+  - \usepackage{subfig}
 ---
@@
 library(tidyverse)
-library(dplyr)
-library(here)
-library(ggplot2)
-library(ggpubr)
+library(palmerpenguins)
 ```
@@
 
-The Great Recession has been a world-wide crisis that reconstructed the economy, and was deep, synchronous and global.
-Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent and that of Japan by more than 8 percent.
-The impact on international trade has been severe that all G20 countries experienced serious reductions in trade[@recession].
-Although it is difficult to compare across countries since policy context, society norms and cultural preferences tend to differ, it is never possible to separately consider the effects on each country even after recovery.
-
-In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas.
-Another impact from depressed economic condition is delayed childbearing, and others to hasten it.
-The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility].
-The significant consequence on fertility rate would make us wonder how severe is America being affected by the great recession.
-However,  by comparing currency devaluations, equity market declines, and rising sovereign bond spreads, the U.S.
-is the third leas affected country[@rank].
-
-Prior research suggests that stress may be linked to riskier sextual activity.
-Financial hardship and poverty can negatively affect cognitive function, and shorten the time horizons over which individuals make cost-benefit determinations.
-Alternatively, fertility may be unaffected by economic conditions if cultural norms related to the context and timing of births are particularly influential.
-For example, non-marital births are more socially consequential for some racial and ethnic groups, and in some geographic areas.
-For some individuals, these social norms may outweigh any economic considerations[@fertility].
-The same consideration of religious belief could also be applied to other countries.
+You can and should cross-reference sections and sub-sections.
+For instance, Section \@ref(data).
+R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
 
@@
 
-We started our analysis by using R [@r], dplyr[@dplyr], tidyverse[@tidyverse]. Graphs are generated using ggpubr[@ggpubr] and ggplot2[@ggplot2].
-```{r, include=FALSE}
-#importing the Birth Rate 1950-2022 of the least affected countries: China, Japan, and the United States from https://www.macrotrends.net/countries/USA/united-states/birth-rate
-#Then modify the 'date' variable of each into Date format for graphs by using as.Date
-china_data <- read_csv(here::here("inputs/data/china-population-2022-02-25.csv"))
-china_data$date <- as.Date(china_data$date)
-str(china_data)
+Our data is of penguins (Figure \@ref(fig:bills)).
 
-japan_data <- read.csv(here::here("inputs/data/japan-population-2022-02-25.csv"))
-japan_data$date <- as.Date(japan_data$date)
-str(japan_data)
-
-us_data <- read_csv(here::here("inputs/data/united-states-population-2022-02-26.csv"))
-us_data$date <- as.Date(us_data$date)
-str(us_data)
-
-sa_data <- read_csv(here::here("inputs/data/south-africa-population-2022-02-26.csv"))
-sa_data$date <- as.Date(sa_data$date)
-str(sa_data)
+```{r bills, fig.cap="Bills of penguins", echo = FALSE}
+ggplot(penguins, aes(x = island, fill = species)) +
+  geom_bar(alpha = 0.8) +
+  scale_fill_manual(values = c("darkorange","purple","cyan4"),
+                    guide = FALSE) +
+  theme_minimal() +
+  facet_wrap(~species, ncol = 1) +
+  coord_flip()
 ```
 
-```{r,echo=FALSE}
-#Generating time series graphs of births per 1000 people for each least affected country
+Talk more about it.
 
-china <- ggplot(china_data,aes(x=date,y= `Births per 1000 People`)) + 
-   labs(title = "China",
-        x = "Year",
-        y = "Births per 1000 People ") +
+Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)
+
+```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
+
+# This needs to be about the random data tha tI sasmpled?
+ggplot(penguins, aes(x = island, fill = species)) +
+  geom_bar(alpha = 0.8) +
+  scale_fill_manual(values = c("darkorange","purple","cyan4"),
+                    guide = FALSE) +
   theme_minimal() +
-  geom_line()
-
-japan <- ggplot(japan_data,aes(x=date,y= Births.per.1000.People)) + 
-   labs(title = "Japan",
-        x = "Year",
-        y = "Births per 1000 People ") +
-  theme_minimal() +
-  geom_line()
-
-us <- ggplot(us_data,aes(x=date,y= `Births per 1000 People`)) + 
-   labs(title = "The United States",
-        x = "Year",
-        y = "Births per 1000 People ") +
-  theme_minimal() +
-  geom_line()
-
-sa <- ggplot(sa_data,aes(x=date,y= `Births per 1000 People`)) + 
-   labs(title = "South Africa",
-        x = "Year",
-        y = "Births per 1000 People ") +
-  theme_minimal() +
-  geom_line()
-
-#Arranging them on the same page
-ggarrange(china, japan, us, sa + rremove("x.text"),
-          ncol = 2, nrow = 2)
+  facet_wrap(~species, ncol = 1) +
+  coord_flip()
 ```
 
-```{r, include=FALSE}
-#importing the Birth Rate 1950-2022 of the most affected countries: Ukraine, Argentina, Hungary, and Poland from https://www.macrotrends.net/countries/USA/united-states/birth-rate
-#Then modify the 'date' variable of each into Date format for graphs by using as.Date
-ukraine_data <- read_csv(here::here("inputs/data/ukraine-population-2022-02-26.csv"))
-ukraine_data$date <- as.Date(ukraine_data$date)
-str(ukraine_data)
-
-argentina_data <- read_csv(here::here("inputs/data/argentina-population-2022-02-26.csv"))
-argentina_data$date <- as.Date(argentina_data$date)
-
-hungary_data <- read_csv(here::here("inputs/data/hungary-population-2022-02-27.csv"))
-hungary_data$date <- as.Date(hungary_data$date)
-
-poland_data <- read_csv(here::here("inputs/data/poland-population-2022-02-26.csv"))
-poland_data$date <- as.Date(poland_data$date)
-
-```
-
-```{r, echo=FALSE}
-#Generating time series graphs of births per 1000 people for each most affected country
-ukraine <- ggplot(ukraine_data,aes(x=date,y= `Births per 1000 People`)) + 
-   labs(title = "Ukraine",
-        x = "Year",
-        y = "Births per 1000 People ") +
-  theme_minimal() +
-  geom_line()
-
-argentina <- ggplot(argentina_data,aes(x=date,y= `Births per 1000 People`)) + 
-   labs(title = "Argentina",
-        x = "Year",
-        y = "Births per 1000 People ") +
-  theme_minimal() +
-  geom_line()
-
-hungary <- ggplot(hungary_data,aes(x=date,y= `Births per 1000 People`)) + 
-   labs(title = "Hungary",
-        x = "Year",
-        y = "Births per 1000 People ") +
-  theme_minimal() +
-  geom_line()
-
-poland <- ggplot(poland_data,aes(x=date,y= `Births per 1000 People`)) + 
-   labs(title = "Poland",
-        x = "Year",
-        y = "Births per 1000 People ") +
-  theme_minimal() +
-  geom_line()
-
-ggarrange(ukraine, argentina, hungary, poland + rremove("x.text"),
-          ncol = 2, nrow = 2)
-```
+Talk way more about it.
 
 # Model
+
+\begin{equation}
+Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
+\end{equation}
+
+Equation \@ref(eq:bayes) seems useful, eh?
+
+Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
+We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
 
@@
 # References
+
+```{r echo = FALSE, fig-sub, fig.cap='Density by State', fig.subcap=c('Same Sex Couple State Density', 'LGBT State Density'), out.width='.49\\linewidth', fig.asp=1, fig.ncol = 2}
+knitr::include_graphics(c("./images/same-sex-couple-state-density.png", "./images/LGBT-state-density.png"))
+```
+# LGBT Section

### 2022-02-27 outputs/paper/paper.Rmd
push

--- 2022-02-27 outputs/paper/paper.Rmd
@@
   - Sijian Wu
-thanks: "Code and data are available at: https://github.com/cuilantao/STA304-Paper-2"
+thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47. Code and data avaliable at https://github.com/cuilantao/STA304-Paper-2"         
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "Between 1980 and the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
+abstract: "This paper is based on Kearney's paper The Puzzle of Falling US Birth Rates since the Great Recession Between 1980.
+^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
 However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years.

### 2022-02-27 outputs/paper/paper.Rmd
Comparison with least and most affected countries

@@
 str(sa_data)
+
 ```
@@
 china <- ggplot(china_data,aes(x=date,y= `Births per 1000 People`)) +
+
    labs(title = "China",
@@
 japan <- ggplot(japan_data,aes(x=date,y= Births.per.1000.People)) +
+
    labs(title = "Japan",
@@
 us <- ggplot(us_data,aes(x=date,y= `Births per 1000 People`)) +
+
    labs(title = "The United States",
@@
 #Then modify the 'date' variable of each into Date format for graphs by using as.Date
+
 ukraine_data <- read_csv(here::here("inputs/data/ukraine-population-2022-02-26.csv"))
@@
 poland_data <- read_csv(here::here("inputs/data/poland-population-2022-02-26.csv"))
+
 poland_data$date <- as.Date(poland_data$date)
@@
 #Generating time series graphs of births per 1000 people for each most affected country
+
 ukraine <- ggplot(ukraine_data,aes(x=date,y= `Births per 1000 People`)) + 

### 2022-02-27 outputs/paper/paper.Rmd
Auto stash before merge of "huakun" and "origin/main"

--- 2022-02-27 outputs/paper/paper.Rmd
@@
 
+## LGBT Section
+
+The percentage of LGBT (and many other genders) has risen in recent years in the US, probably related to propaganda and political correctness.
+
+Since there is not much data for other genders, and LGBT are the majorities, we will focus on LGBT population and the potential effect they bring to birth rate.
+
+LGBT is the acronym of the following 4 genders, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender.
+By definition, these genders by definition should have a much lower birth rate than heterosexual.
+Thus, the population changes in these population are very likely to make an impact on the birth rate of the US.
+
+```{r echo = FALSE, fig-gender-raising-children-distribution, fig.cap='Gender Distribution and Percentage Raising Children Distribution', fig.subcap=c('LGBT Gender Distribution', 'LGBT Percentage Raising Children', 'Same Sex Couple Gender Distribution'), out.width='.33\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c("./images/LGBT-Gender-distribution.png",
+  "./images/LGBT-percentage-raising-children.png",
+  "./images/same-sex-couple-gender-distribution.png"))
+```
+
+Figure \@ref(fig:fig-gender-raising-children-distribution) shows the gender distribution of LGBT and Same Sex Couple as well as the percentage raising children.
+Figure \@ref(fig:fig-gender-raising-children-distribution) (a) and Figure \@ref(fig:fig-gender-raising-children-distribution) (c) clearly shows that more females are non-heterosexual.
+This should make a even larger impact on birth rate as only females are able to give births (in normal situations).
+
+Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children.
+71% of the LGBT population don't raise children.
+With these information, we can conclude that the growth of LGBT population is able to cause a negative change in birth rate.
+
+```{r echo = FALSE, fig-density-state, fig.cap='Density by State', fig.subcap=c('Same Sex Couple State Density', 'LGBT State Density', 'Birth Rate Distribution by State'), out.width='.49\\linewidth', fig.asp=1, fig.ncol = 2}
+knitr::include_graphics(c("./images/same-sex-couple-state-density.png", "./images/LGBT-state-density.png", "./images/state-birth-rate-distribution.png"))
+```
+
+Figure \@ref(fig:fig-density-state) shows the population density of Same Sex Couple and LGBT by States, and birth rate distribution by State.
+Darker color means higher density or higher negative change in birth rate.
+From these plots we can see that the states with higher LGBT density and same sex couples generally have higher negative change in birth rate.
+We cannot conclude whether LGBT population has a large impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.
+
+[@cite_original]
+
+```{r echo = FALSE, fig-race-distribution, fig.cap='Race Distribution', fig.subcap=c('Same Sex Couple Race Distribution', 'LGBT Race Distribution'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2, fig.align="center"}
+knitr::include_graphics(c("./images/same-sex-couple-race-distribution.png", "./images/LGBT-race-distribution.png"))
+```
+
+```{r echo = FALSE, fig-age-distribution, fig.cap='LGBT and Same Sex Couple Age Distribution', fig.subcap=c('Same Sex Couple Race Distribution', 'LGBT Race Distribution'), out.width='.40\\linewidth', fig.asp=1, fig.ncol = 2, fig.align="center"}
+knitr::include_graphics(c("./images/same-sex-couple-age-distribution.png", "./images/LGBT-age-distribution.png"))
+```
+
+```{r echo = FALSE, fig-LGBT-socioeconomic-indicators, fig.cap='LGBT Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
+knitr::include_graphics("./images/LGBT-socioeconomic-indicators.png")
+```
+
 ## Third discussion point

### 2022-02-28 outputs/paper/paper.Rmd

@@
 By definition, these genders by definition should have a much lower birth rate than heterosexual.
+Then change in birth rate and change in LGBT population should have a negative propotional relationship (i.e.
+Growth in LGBT population leads to lower birth rate).
 Thus, the population changes in these population are very likely to make an impact on the birth rate of the US.
+
+If we assume the simpliest case where we ignore race, gender distribution (equal percentage of male and female) and all other factors, and assume that no LGBT or same sex couple population give birth to a child, then there would be a perfect correlation (perfectly proportional) between LGBT population change and birth rate (i.e.
+a $x\%$ increase in LGBT population could lead to $x\%$ decrease in birth rate).
+Let's define $\phi = \frac{\Delta Birth Rate}{\Delta LGBTPopulation}$ and call $\phi$ the BirthRate-LGBT coefficient for convenience.
+If we take other factors into account, the impact to birth rate could be even larger (i.e.
+$\phi<-1$).
 
@@
 Figure \@ref(fig:fig-gender-raising-children-distribution) (a) and Figure \@ref(fig:fig-gender-raising-children-distribution) (c) clearly shows that more females are non-heterosexual.
-This should make a even larger impact on birth rate as only females are able to give births (in normal situations).
+This should make a even larger impact ($\phi<-1$) on birth rate as only females are able to give births (in normal situations).
 
-Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children.
-71% of the LGBT population don't raise children.
-With these information, we can conclude that the growth of LGBT population is able to cause a negative change in birth rate.
+Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children. 71% of the LGBT population don't raise children.
 
@@
 
-[@cite_original]
-
 ```{r echo = FALSE, fig-race-distribution, fig.cap='Race Distribution', fig.subcap=c('Same Sex Couple Race Distribution', 'LGBT Race Distribution'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2, fig.align="center"}
@@
 ```
+
+From \@ref(fig:fig-race-distribution) we can see that White population takes up the majority of the population of LGBT and Same Sex Couple.
+White is also the majority of the population of the US.
+Thus, larger LGBT population could lead to a even larger decrease in birth rate ($\phi<-1$).
 
@@
 
-```{r echo = FALSE, fig-LGBT-socioeconomic-indicators, fig.cap='LGBT Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
+From \@ref(fig:fig-age-distribution) (a) we can see that most same sex couples fall in the age group 25-64.
+This period overlaps with the majority of the childbearing age.
+From \@ref(fig:fig-age-distribution) (b), we can see that LGBT population decreases as age increases starting from the age or 18, which is the beginning of women's childbearing age.
+These 2 evidences also growth of LGBT population could lead to a larger decrease in birth rate ($\phi < -1$).
+
+With these information, we can conclude that the growth of LGBT population is able to cause a negative change in birth rate.
+
+```{r echo = FALSE, fig-LGBT-socioeconomic-indicators, fig.cap='Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
 knitr::include_graphics("./images/LGBT-socioeconomic-indicators.png")
@@
 
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+Figure \@ref(fig:fig-LGBT-socioeconomic-indicators) shows some Socioeconomic Indicators classified by sexual orientation (LGBT or not), such as employment status, income level education level, and whether or not being insured and food insecure.
+Darker color represents LGBT, lighter color represents non-LGBT.
+From the plots, we can extract some socioeconomic indicators of LGBT.
+Comparing to non-LGBT population, LGBT are more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is lower than that of the percentage of non-LGBT population).
 

### 2022-02-28 outputs/paper/paper.Rmd
Update LGBT analysis

@@
 
-The percentage of LGBT (and many other genders) has risen in recent years in the US, probably related to propaganda and political correctness.
+The percentage of LGBT (and many other gender identities) population has risen in recent years in the US, probably related to propaganda and political correctness.
 
@@
 
-LGBT is the acronym of the following 4 genders, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender.
-By definition, these genders by definition should have a much lower birth rate than heterosexual.
+LGBT is the acronym of the following 4 gender identities, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender.
+By definition, these genders should have a much lower birth rate than heterosexual.
 Then change in birth rate and change in LGBT population should have a negative propotional relationship (i.e.
@@
 This should make a even larger impact ($\phi<-1$) on birth rate as only females are able to give births (in normal situations).
-
+went
 Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children. 71% of the LGBT population don't raise children.

### 2022-02-28 outputs/paper/paper.Rmd
Update LGBT analysis

@@
 LGBT is the acronym of the following 4 gender identities, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender.
-By definition, these genders should have a much lower birth rate than heterosexual.
-Then change in birth rate and change in LGBT population should have a negative propotional relationship (i.e.
-Growth in LGBT population leads to lower birth rate).
+By definition, these genders should have a much lower birth rate than that of the heterosexual as normal births of human must depend on a male and a female.
+Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children. 71% of the LGBT population don't raise children. 
+Then changes in birth rate and changes in LGBT population should have a negative proportional relationship (i.e. Growth in LGBT population leads to lower birth rate).
 Thus, the population changes in these population are very likely to make an impact on the birth rate of the US.
 
-If we assume the simpliest case where we ignore race, gender distribution (equal percentage of male and female) and all other factors, and assume that no LGBT or same sex couple population give birth to a child, then there would be a perfect correlation (perfectly proportional) between LGBT population change and birth rate (i.e.
-a $x\%$ increase in LGBT population could lead to $x\%$ decrease in birth rate).
+If we assume the simpliest scenario where we ignore race, gender distribution (equal percentage of male and female) and all other factors, and assume that no LGBT or same-sex-couple population give birth to a child, then there would be a perfect correlation (perfectly proportional) between LGBT population change and birth rate (i.e.
+an $x\%$ increase in LGBT population could lead to an $x\%$ decrease in birth rate).
 Let's define $\phi = \frac{\Delta Birth Rate}{\Delta LGBTPopulation}$ and call $\phi$ the BirthRate-LGBT coefficient for convenience.
+Basically, lower BirthRate-LGBT coefficient means lower birth rate while holding LGBT population constant.
 If we take other factors into account, the impact to birth rate could be even larger (i.e.
@@
 
-Figure \@ref(fig:fig-gender-raising-children-distribution) shows the gender distribution of LGBT and Same Sex Couple as well as the percentage raising children.
-Figure \@ref(fig:fig-gender-raising-children-distribution) (a) and Figure \@ref(fig:fig-gender-raising-children-distribution) (c) clearly shows that more females are non-heterosexual.
-This should make a even larger impact ($\phi<-1$) on birth rate as only females are able to give births (in normal situations).
-went
-Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children. 71% of the LGBT population don't raise children.
+Figure \@ref(fig:fig-gender-raising-children-distribution) shows the gender distribution of LGBT and Same Sex Couple as well as well as the percentage raising children.
+Figure \@ref(fig:fig-gender-raising-children-distribution) (a) and Figure \@ref(fig:fig-gender-raising-children-distribution) (c) clearly shows that more females are non-heterosexual than male are.
+This could cause a lower BirthRate-LGBT coefficient on birth rate as only females are able to give births (in normal situations).
 
@@
 
-Figure \@ref(fig:fig-density-state) shows the population density of Same Sex Couple and LGBT by States, and birth rate distribution by State.
-Darker color means higher density or higher negative change in birth rate.
-From these plots we can see that the states with higher LGBT density and same sex couples generally have higher negative change in birth rate.
-We cannot conclude whether LGBT population has a large impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.
+Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US.
+Figure \@ref(fig:fig-density-state) [@cite_original] (c) shows the birth rate distribution by State in the US.
+Darker color means higher density or larger negative change in birth rate.
+From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate.
+States on the west generally have darker color in all three plots.
+We cannot conclude whether LGBT population has a large or direct impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.
 
@@
 White is also the majority of the population of the US.
-Thus, larger LGBT population could lead to a even larger decrease in birth rate ($\phi<-1$).
+Thus, larger LGBT population could lead to a even larger decrease in birth rate ($\phi<-1$, lower BirthRate-LGBT coefficient).
 
@@
 From \@ref(fig:fig-age-distribution) (b), we can see that LGBT population decreases as age increases starting from the age or 18, which is the beginning of women's childbearing age.
-These 2 evidences also growth of LGBT population could lead to a larger decrease in birth rate ($\phi < -1$).
+Around 90% of the LGBT population have their age between 18 and 64, which covers almost all of the child bearing ages.
+These 2 evidences also indicate that, growth of LGBT population could lead to a larger decrease in birth rate ($\phi < -1$, lower BirthRate-LGBT coefficient).
 

### 2022-02-28 outputs/paper/paper.Rmd
country intro & data description

@@
 
+As mentioned by @cite_original, the basic approach to many economic problem is maximizing utility.
+This approach recognized that children bring people utility but also with associated costs including time and money.
+This approach to modeling the decision to have a child leads to standard predictions of price and income effects.
+Direct consideration of costs could be like housing and medical care.
+Other times the price effect is indirect, like the opportunity cost of a woman’s time.
+Holding prices and quality constant, an increase in income will lead people to choose to have more children.
+This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per capita GDP observed over time and place[@cite_original].
+However, this is enough for us to consider whether the birth rates are world-wide affected in the same negative direction after the Great Recession.
+
 # Data
@@
 We started our analysis by using R [@r], dplyr[@dplyr], tidyverse[@tidyverse]. Graphs are generated using ggpubr[@ggpubr] and ggplot2[@ggplot2].
+
+We generated data from U.S.
+Birth Rate 1950-2022[@birthrate].
+We selected top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
 ```{r, include=FALSE}

### 2022-02-28 outputs/paper/paper.Rmd
pulling change

--- 2022-02-28 outputs/paper/paper.Rmd
@@
   - Sijian Wu
-thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47. Code and data avaliable at https://github.com/cuilantao/STA304-Paper-2"         
+thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47.
+Code and data avaliable at https://github.com/cuilantao/Investigation-of-falling-US-Birth-Rates-since-the-great-recession"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years.
-In this paper, we aim to find out the reason behind the decline, all of the data are collected from the CDC Vital Statistics Births Reports.
-Our findings can help to understand the reason behind the drop.
-"
+In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries.
+We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect"
 output:
@@
 The impact on international trade has been severe that all G20 countries experienced serious reductions in trade[@recession].
-Although it is difficult to compare across countries since policy context, society norms and cultural preferences tend to differ, it is never possible to separately consider the effects on each country even after recovery.
+Apart from economic impact, the great recession also had a negative impact on country's birth rate, Figure 1 shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession.
+
+```{r, fig.cap="US Birth Rate in the past 40 years", echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=4}
+US_birth_date_from_1985_to_2020 <- read_csv(here::here("inputs/data/Birthrate.csv"))
+
+ggplot(data=US_birth_date_from_1985_to_2020, aes(x=`year`, y=`brate_all`, group=1)) +
+  geom_line()+
+  geom_point() +
+  ylim(50, 80) +
+  xlab("Year") + ylab("Births per 1,000 women age 15-44")
+
+```
 
@@
 
-Prior research suggests that stress may be linked to riskier sextual activity.
+Prior research suggests that stress may be linked to riskier sexual activity.
 Financial hardship and poverty can negatively affect cognitive function, and shorten the time horizons over which individuals make cost-benefit determinations.
@@
 
+Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started at the 1920s and had huge progress after 2000s.
+A lot of laws that prohibited homosexual relationship were lifted and homosexual couples can get married legally and adopt kids, there are also propaganda that encourages homosexuality.
+
+In this paper, we will investigate the effect of the Great Recession on the birth rate in different countries, and we will also explore the effect of LGBT movement on birth rate.
+
 # Data
@@
 We started our analysis by using R [@r], dplyr[@dplyr], tidyverse[@tidyverse]. Graphs are generated using ggpubr[@ggpubr] and ggplot2[@ggplot2].
+
 ```{r, include=FALSE}

### 2022-02-28 outputs/paper/paper.Rmd

@@
 
+As mentioned by @cite_original, the basic approach to many economic problem is maximizing utility.
+This approach recognized that children bring people utility but also with associated costs including time and money.
+This approach to modeling the decision to have a child leads to standard predictions of price and income effects.
+Direct consideration of costs could be like housing and medical care.
+Other times the price effect is indirect, like the opportunity cost of a woman’s time.
+Holding prices and quality constant, an increase in income will lead people to choose to have more children.
+This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per capita GDP observed over time and place[@cite_original].
+However, this is enough for us to consider whether the birth rates are world-wide affected in the same negative direction after the Great Recession.
+
 Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started at the 1920s and had huge progress after 2000s.
@@
 We started our analysis by using R [@r], dplyr[@dplyr], tidyverse[@tidyverse]. Graphs are generated using ggpubr[@ggpubr] and ggplot2[@ggplot2].
+
+## Data Source
+We generated data from U.S.
+Birth Rate 1950-2022[@birthrate].
+We selected top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
 

### 2022-02-28 outputs/paper/paper.Rmd
Auto stash before merge of "main" and "origin/main"

@@
 library(ggpubr)
+library(haven)
+library(tidyr)
+library(sqldf)
 ```
@@
 
+```{r echo = FALSE, fig-cohort, fig.cap='Children Ever Born by Mother\'s Age and Birth Cohort', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
+knitr::include_graphics("./images/cohort.png")
+```
+
+```{r cohort-plot, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.cap="Children Ever Born by Mother Age and Birth Cohort", fig.height=4, fig.width=8}
+age_comp_pop <- read_dta(here::here("inputs/data/agecomp-seer.dta"))
+age_comp_pop <- subset(age_comp_pop, select = -c(fem1519, fem2034, fem3544, stname))
+age_comp_pop_grouped <- group_by(age_comp_pop, year) %>% summarise(across(everything(), sum))
+age_comp_pop_grouped_start_with_fem <-
+  select(age_comp_pop_grouped, starts_with("fem"))
+fem_col_name <- colnames(age_comp_pop_grouped_start_with_fem)
+age_comp_pop_long <- gather(age_comp_pop_grouped, mage, fem, fem_col_name)
+age_comp_pop_long$mage <- gsub("^.{0,3}", "", age_comp_pop_long$mage)
+age_comp_pop_long$mage <- as.numeric(as.character(age_comp_pop_long$mage))
+age_comp_pop_long$cohort <- with(age_comp_pop_long, year - mage)
+age_comp_pop_long$cohort2 <- ''
+
+age_comp_pop_long$cohort2[age_comp_pop_long$cohort >= 1968 & age_comp_pop_long$cohort <= 1972] <- 1
+age_comp_pop_long$cohort2[age_comp_pop_long$cohort >= 1973 & age_comp_pop_long$cohort <= 1977] <- 2
+age_comp_pop_long$cohort2[age_comp_pop_long$cohort >= 1978 & age_comp_pop_long$cohort <= 1982] <- 3
+age_comp_pop_long$cohort2[age_comp_pop_long$cohort >= 1983 & age_comp_pop_long$cohort <= 1987] <- 4
+age_comp_pop_long$cohort2[age_comp_pop_long$cohort >= 1988 & age_comp_pop_long$cohort <= 1992] <- 5
+age_comp_pop_long$cohort2[age_comp_pop_long$cohort >= 1993 & age_comp_pop_long$cohort <= 1997] <- 6
+age_comp_pop_long <- subset(age_comp_pop_long, cohort2 != '')
+
+names(age_comp_pop_long)[names(age_comp_pop_long) == 'fem'] <- 'pop'
+age_comp_pop_long_group <- group_by(age_comp_pop_long, cohort2, mage) %>% summarise(pop_sum = sum(pop))
+cohort_analysis <- read_dta(here::here("inputs/data/nchs_cohort_analysis.dta"))
+merged_data <- merge(cohort_analysis, age_comp_pop_long_group, by = c("cohort2", "mage"))
+merged_data$brate <- with(merged_data, numbirth/pop_sum*1000)
+data_with_only_cohort_and_brate <- subset(merged_data, select = c(cohort2, brate, mage))
+data_with_only_cohort_and_brate <- data_with_only_cohort_and_brate[order(data_with_only_cohort_and_brate$cohort2, data_with_only_cohort_and_brate$mage),]
+data_with_only_cohort_and_brate$brate <- ave(data_with_only_cohort_and_brate$brate, data_with_only_cohort_and_brate$cohort2, FUN=cumsum)
+data_with_only_cohort_and_brate_grouped <- group_by(data_with_only_cohort_and_brate, cohort2, mage) %>% summarise(cum_brate = sum(brate) / 1000)
+names(data_with_only_cohort_and_brate_grouped)[names(data_with_only_cohort_and_brate_grouped) == 'cohort2'] <- 'cohort'
+final_data_wide_format <- spread(data_with_only_cohort_and_brate_grouped, key = cohort, value = cum_brate,  sep = '')
+
+data_with_cohort1 <- drop_na(final_data_wide_format, cohort1)
+data_with_cohort2 <- drop_na(final_data_wide_format, cohort2)
+data_with_cohort3 <- drop_na(final_data_wide_format, cohort3)
+data_with_cohort4 <- drop_na(final_data_wide_format, cohort4)
+data_with_cohort5 <- drop_na(final_data_wide_format, cohort5)
+data_with_cohort6 <- drop_na(final_data_wide_format, cohort6)
+
+ggplot() +
+  geom_line(data = data_with_cohort1, aes(x = mage, y = cohort1, color = "1968-1972"))+
+  geom_line(data = data_with_cohort2, aes(x = mage, y = cohort2, color = "1973–1977")) +
+  geom_line(data = data_with_cohort3, aes(x = mage, y = cohort3, color = "1977–1982")) +
+  geom_line(data = data_with_cohort4, aes(x = mage, y = cohort4, color = "1983–1987")) +
+  geom_line(data = data_with_cohort5, aes(x = mage, y = cohort5, color = "1988–1992")) +
+  geom_line(data = data_with_cohort6, aes(x = mage, y = cohort6, color = "1993–1997")) +
+  scale_color_manual(name = "Birth Cohorts", 
+                     values = c("1968-1972" = "green", 
+                                "1973–1977" = "blue", 
+                                "1977–1982" = "orange", 
+                                "1983–1987" = "gray",
+                                "1988–1992" = "yellow4",
+                                "1993–1997" = "lightblue"
+                                )) + theme(legend.position = c(0.9, 0.5)) +
+  xlab("Mother's Age") + ylab("Children ever born") + 
+  ggtitle("Children Ever Born by Mother’s Age and Birth Cohort")
+```
+
+Research has shown that cohort effects may have contributed to the declines of birth rate in the US[@cite_original].
+"Period effects reflect changes that affect everyone at a point in time, whereas cohort effects reflect changes across people born or raised in different years.
+"[@cite_original].
+We reproduced the plot from [@cite_original] in Figure \@ref(fig:cohort-plot).
+
 \newpage

### 2022-02-28 outputs/paper/paper.Rmd
fixed conflict, add abstract

--- 2022-02-28 outputs/paper/paper.Rmd
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "This paper is based on Kearney's paper The Puzzle of Falling US Birth Rates since the Great Recession Between 1980.
+abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\".
 ^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
@@
 The impact on international trade has been severe that all G20 countries experienced serious reductions in trade[@recession].
-Apart from economic impact, the great recession also had a negative impact on country's birth rate, Figure 1 shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession.
-
-```{r, fig.cap="US Birth Rate in the past 40 years", echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=4}
-US_birth_date_from_1985_to_2020 <- read_csv(here::here("inputs/data/Birthrate.csv"))
-
-ggplot(data=US_birth_date_from_1985_to_2020, aes(x=`year`, y=`brate_all`, group=1)) +
-  geom_line()+
-  geom_point() +
-  ylim(50, 80) +
-  xlab("Year") + ylab("Births per 1,000 women age 15-44")
-
-```
+Apart from economic impact, the great recession also had a negative impact on country's birth rate.
 
@@
 However,  by comparing currency devaluations, equity market declines, and rising sovereign bond spreads, the U.S.
-is the third leas affected country[@rank].
+is the third less affected country[@rank].
 
@@
 Holding prices and quality constant, an increase in income will lead people to choose to have more children.
-This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per capita GDP observed over time and place[@cite_original].
+This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per-capita GDP observed over time and place[@cite_original].
 However, this is enough for us to consider whether the birth rates are world-wide affected in the same negative direction after the Great Recession.
@@
 ## Data Source
-We generated data from U.S.
+Part of our data are from the NCHS(National Center for Health Statistics)[@nvss], in particular, we acquired US birth rate data for the past 40 years, Figure \@ref(fig:fig-past-40-years) shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession.
+We also generated data from U.S.
 Birth Rate 1950-2022[@birthrate].
 We selected top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
+TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT
+
+## Data Characteristics
+
+```{r, fig-past-40-years, fig.cap="US Birth Rate in the past 40 years", echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=4}
+US_birth_date_from_1985_to_2020 <- read_csv(here::here("inputs/data/Birthrate.csv"))
+
+ggplot(data=US_birth_date_from_1985_to_2020, aes(x=`year`, y=`brate_all`, group=1)) +
+  geom_line()+
+  geom_point() +
+  ylim(50, 80) +
+  xlab("Year") + ylab("Births per 1,000 women age 15-44")
+
+```
 
@@
 
-```{r,echo=FALSE}
+```{r, include=FALSE}
+#importing the Birth Rate 1950-2022 of the most affected countries: Ukraine, Argentina, Hungary, and Poland from https://www.macrotrends.net/countries/USA/united-states/birth-rate
+#Then modify the 'date' variable of each into Date format for graphs by using as.Date
+
+ukraine_data <- read_csv(here::here("inputs/data/ukraine-population-2022-02-26.csv"))
+ukraine_data$date <- as.Date(ukraine_data$date)
+str(ukraine_data)
+
+argentina_data <- read_csv(here::here("inputs/data/argentina-population-2022-02-26.csv"))
+argentina_data$date <- as.Date(argentina_data$date)
+
+hungary_data <- read_csv(here::here("inputs/data/hungary-population-2022-02-27.csv"))
+hungary_data$date <- as.Date(hungary_data$date)
+
+poland_data <- read_csv(here::here("inputs/data/poland-population-2022-02-26.csv"))
+
+poland_data$date <- as.Date(poland_data$date)
+
+```
+
+# Results
+
+## Global Comparison of Fertility Rate Under The Great Recession
+It shows that although the United States is at the epicenter of the global economic crisis, it is one of the countries least affected by the financial fallout.
+Large industrialized nations like the United States, Japan, and Germany have benefited from increasing global demand for relatively stable economies in which to invest.
+Instead, it is several developing countries, notably those with vulnerable capital accounts and weak macroeconomic fundamentals, that are experiencing severe economic downturns disproportionate to their roles in the crisis[@rank].
+
+```{r, least-affected, echo=FALSE, fig.cap="Births per 1000 People of Top Least Affected Countries"}
 #Generating time series graphs of births per 1000 people for each least affected country
@@
 
-```{r, include=FALSE}
-#importing the Birth Rate 1950-2022 of the most affected countries: Ukraine, Argentina, Hungary, and Poland from https://www.macrotrends.net/countries/USA/united-states/birth-rate
-#Then modify the 'date' variable of each into Date format for graphs by using as.Date
+Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with prediction for the future.
+The births rates of China and South Africa were obviously higher than the U.S.
+and Japan.
+However, around the year of recession, the birth rate of China and Japan dropped even harder than the U.S.
+on the absolute level that are lower than 10 births per 1000 people.
 
-ukraine_data <- read_csv(here::here("inputs/data/ukraine-population-2022-02-26.csv"))
-ukraine_data$date <- as.Date(ukraine_data$date)
-str(ukraine_data)
-
-argentina_data <- read_csv(here::here("inputs/data/argentina-population-2022-02-26.csv"))
-argentina_data$date <- as.Date(argentina_data$date)
-
-hungary_data <- read_csv(here::here("inputs/data/hungary-population-2022-02-27.csv"))
-hungary_data$date <- as.Date(hungary_data$date)
-
-poland_data <- read_csv(here::here("inputs/data/poland-population-2022-02-26.csv"))
-
-poland_data$date <- as.Date(poland_data$date)
-
-```
-
-```{r, echo=FALSE}
+```{r, most-affected, echo=FALSE, fig.cap="Births per 1000 People of Top Most Affected Countries"}
 #Generating time series graphs of births per 1000 people for each most affected country
@@
 
-# Model
+Figure \@ref(fig:most-affected) shows the changes of births per 1000 people of Ukraine, Argentina, Hungary, Poland since 1950 with prediction for the future.
+All these most affected countries have lower initial births rate than those least affected countries.
+However, the levels of births per 1000 people were similar to those least affected countries around recession except Argentina.
+Ukraine, Hungary, and Poland also showed a potential increase after recession but continued to decline afterwards.
 
-# Results
+We do observe declinations of world-wide birth rate despite of the impact level from recession.
+From the economic perspective, the countries who are least affected tend to have a stronger resistance to recession but not necessarily developed countries.
+While the most affected countries are mostly developing countries, they experienced a more significant plummet as percentage.
+It is also intriguing that the long-term births per 1000 people tend to converge to 10 for all countries mentioned above.
+However, the countries have higher resistance tend to delay the convergence which may derived from stronger economic status.
+Overall, the Great Recession was a turning point not only for the U.S.
+but also the whole world in the context of birth rate.
+The Great Recession may not be the direct cause of the world-wide declination in birth rates but it indeed provided an opportunity and provoked the desire for people to reconsider personal priorities with changed perspective of the society.
 
-# Discussion
-
-## First discussion point
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## LGBT Section
+## Effect of LGBT movement
 

### 2022-02-28 scripts/replicate_graph_02.R

@@ -5,21 +5,25 @@ library(here)
 library(ggplot2)
 library(haven)
 
+# Import the dta
 birth_pop <- read_dta(here::here("inputs/data/nchs_births_pop_1990_2019.dta"))
+# Import the dta
 age_race_pop <- read_dta(here::here("inputs/data/age_race_comp_seer.dta"))
-
+# Mege two dta together
 merged_data <- merge(birth_pop, age_race_pop, by = c("stname", "year"))
 
 # Define new column
 merged_data$pop2044 <- with(merged_data, `pop2024.y` + `pop2534.y` + `pop3544.y`)
-
+# use a group by to get the sum
 merged_data_grouped_by <- group_by(merged_data, year) %>% summarise(fb_sum = sum(numbirth_firstbirth), sb_sum = sum(numbirth_secondbirth), tb_sum = sum(numbirth_thirdbirth), fbp_sum = sum(numbirth_fourthbirth), pop_sum = sum(pop1544))
 
+#define new columns based on existing columns
 merged_data_grouped_by$brith_rate_fb <- with(merged_data_grouped_by, fb_sum / `pop_sum` * 1000)
 merged_data_grouped_by$brith_rate_sb <- with(merged_data_grouped_by, sb_sum / `pop_sum` * 1000)
 merged_data_grouped_by$brith_rate_tb <- with(merged_data_grouped_by, tb_sum / `pop_sum` * 1000)
 merged_data_grouped_by$brith_rate_fbp <- with(merged_data_grouped_by, fbp_sum / `pop_sum` * 1000)
 
+# merge the plot and add legends
 ggplot(merged_data_grouped_by, aes(x=year)) + 
  geom_line(aes(y = brith_rate_fb, color = "First Birth")) + 
   geom_line(aes(y = brith_rate_sb, color="Second Birth")) +

### 2022-02-28 outputs/paper/paper.Rmd
Add discussion about cohort effect.

--- 2022-02-28 outputs/paper/paper.Rmd
@@
 
-```{r echo = FALSE, fig-cohort, fig.cap='Children Ever Born by Mother\'s Age and Birth Cohort', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
-knitr::include_graphics("./images/cohort.png")
-```
-
 ```{r cohort-plot, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.cap="Children Ever Born by Mother Age and Birth Cohort", fig.height=4, fig.width=8}
@@
 "[@cite_original].
-We reproduced the plot from [@cite_original] in Figure \@ref(fig:cohort-plot).
+
+We reproduced the plot showing Fertility Rate vs Mother's age from [@cite_original] in Figure \@ref(fig:cohort-plot).
+This plot is basically showing the fertility rate of women in different ages across 6 generations.
+It's quite obvious that the 3 oldest cohorts (generations) have similar and higher fertility rates comparing to that of the 3 younger cohorts (generations).The 3 younger cohorts also have lower and lower fertility rate as their birthday moves forward (younger generation -> lower fertility rate).
+
+```{r fig-parity, fig.cap="Parity (ages 15-44)", echo=FALSE, fig.height=4, fig.height=3}
+birth_pop <- read_dta(here::here("inputs/data/nchs_births_pop_1990_2019.dta"))
+age_race_pop <- read_dta(here::here("inputs/data/age_race_comp_seer.dta"))
+
+merged_data <- merge(birth_pop, age_race_pop, by = c("stname", "year"))
+
+# Define new column
+merged_data$pop2044 <- with(merged_data, `pop2024.y` + `pop2534.y` + `pop3544.y`)
+
+merged_data_grouped_by <- group_by(merged_data, year) %>% summarise(fb_sum = sum(numbirth_firstbirth), sb_sum = sum(numbirth_secondbirth), tb_sum = sum(numbirth_thirdbirth), fbp_sum = sum(numbirth_fourthbirth), pop_sum = sum(pop1544))
+
+merged_data_grouped_by$brith_rate_fb <- with(merged_data_grouped_by, fb_sum / `pop_sum` * 1000)
+merged_data_grouped_by$brith_rate_sb <- with(merged_data_grouped_by, sb_sum / `pop_sum` * 1000)
+merged_data_grouped_by$brith_rate_tb <- with(merged_data_grouped_by, tb_sum / `pop_sum` * 1000)
+merged_data_grouped_by$brith_rate_fbp <- with(merged_data_grouped_by, fbp_sum / `pop_sum` * 1000)
+
+ggplot(merged_data_grouped_by, aes(x=year)) + 
+ geom_line(aes(y = brith_rate_fb, color = "First Birth")) + 
+  geom_line(aes(y = brith_rate_sb, color="Second Birth")) +
+  geom_line(aes(y = brith_rate_tb, color="Third Birth")) +
+  geom_line(aes(y = brith_rate_fbp, color="Fourth Birth")) +
+  scale_color_manual(name = "Parity (ages 15-44)", 
+                     values = c("First Birth" = "steelblue", 
+                                "Second Birth" = "orange", 
+                                "Third Birth" = "gray", 
+                                "Fourth Birth" = "yellow")) + 
+  xlab("Year") + ylab("Birth Rate (births per 1000 women)") + ggtitle("Parity (ages 15-44)")
+```
+
+Figure \@ref(fig:fig-parity) is a replication of a plot showing trends in births by "parity" (referring to the number of children for a given woman) from [@cite_original].
+First and second births declined a lot since 2007, but the trend lines for third and higher order births are comparatively flat.
+The women giving their 3rd and 4th births are most likely an older cohort than the cohorts giving their 1st and 2nd births.
+Again, this is another evidence showing the impact brought by cohort effects.
+
+```{r echo = FALSE, fig-lgbt-trend-by-year, fig.cap='Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
+knitr::include_graphics("./images/US-LGBT-percent-trend.png")
+```
+
+```{r echo = FALSE, fig-lgbt-by-generation, fig.cap='Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
+knitr::include_graphics("./images/LGBT-percentage-by-generation.png")
+```
+
+Figure \@ref(fig:fig-lgbt-trend-by-year) shows the trend of LGBT population percentage in total population.
+There is an obvious increasing trend (i.e. higher percentage of the US population are becoming LGBT). 
+Figure \@ref(fig:fig-lgbt-by-generation) shows the percentage of LGBT population trend in different generations.
+We can also see an increasing trend of LGBT population percentage (younger generations are more and more likely to become LGBT).
+This provide more evidence that cohort effects may contribute to the delines of birth rate in the US in recent years.
 

### 2022-02-28 scripts/replicate_graph_02.R

@@ -30,7 +30,15 @@ ggplot(merged_data_grouped_by, aes(x=year)) +
                                 "Second Birth" = "orange", 
                                 "Third Birth" = "gray", 
                                 "Fourth Birth" = "yellow")) + 
-  xlab("Year") + ylab("Parity") + ggtitle("Parity (ages 15-44)")
+  xlab("Year") + ylab("Birth Rate (births per 1000 women)") + ggtitle("Parity (ages 15-44)")
+
+
+
+
+
+
+
+
 
 
 

### 2022-02-28 outputs/paper/paper.Rmd
fixed conflict, add abstract

--- 2022-02-28 outputs/paper/paper.Rmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "What Happened to The Birth Rate?"
+subtitle: "The Great Recession and the LGBT Movement"
 author: 
@@
 abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\".
-^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
+^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
 However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years.
@@
 
-The Great Recession has been a world-wide crisis that reconstructed the economy, and was deep, synchronous and global.
-Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent and that of Japan by more than 8 percent.
+The Great Recession has been a worldwide crisis that reconstructed the economy and was deep, synchronous and global.
+Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent, and that of Japan by more than 8 percent.
 The impact on international trade has been severe that all G20 countries experienced serious reductions in trade[@recession].
-Apart from economic impact, the great recession also had a negative impact on country's birth rate.
+Apart from the economic impact, the great recession also had a negative impact on the country's birth rate.
 
 In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas.
-Another impact from depressed economic condition is delayed childbearing, and others to hasten it.
+Another impact from the depressed economic condition is delayed childbearing, and others to hasten it.
 The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility].
@@
 
-As mentioned by @cite_original, the basic approach to many economic problem is maximizing utility.
-This approach recognized that children bring people utility but also with associated costs including time and money.
+As mentioned by @cite_original, the basic approach to many economic problems is maximizing utility.
+This approach recognized that children bring people utility but also comes with associated costs including time and money.
 This approach to modeling the decision to have a child leads to standard predictions of price and income effects.
 Direct consideration of costs could be like housing and medical care.
-Other times the price effect is indirect, like the opportunity cost of a woman’s time.
+Other times the price effect is indirect, as the opportunity cost of a woman’s time.
 Holding prices and quality constant, an increase in income will lead people to choose to have more children.
 This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per-capita GDP observed over time and place[@cite_original].
-However, this is enough for us to consider whether the birth rates are world-wide affected in the same negative direction after the Great Recession.
+However, this is enough for us to consider whether the birth rates are worldwide affected in the same negative direction after the Great Recession.
 
-Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started at the 1920s and had huge progress after 2000s.
-A lot of laws that prohibited homosexual relationship were lifted and homosexual couples can get married legally and adopt kids, there are also propaganda that encourages homosexuality.
+Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started in the 1920s and had huge progress after the 2000s.
+A lot of laws that prohibited homosexual relationships were lifted and homosexual couples can get married legally and adopt kids, there is also propaganda that encourages homosexuality.
 
-In this paper, we will investigate the effect of the Great Recession on the birth rate in different countries, and we will also explore the effect of LGBT movement on birth rate.
+In this paper, we will investigate the effect of the Great Recession on the birth rate in different countries, and we will also explore the effect of the LGBT movement on the birth rate.
 
@@
 ## Data Source
-Part of our data are from the NCHS(National Center for Health Statistics)[@nvss], in particular, we acquired US birth rate data for the past 40 years, Figure \@ref(fig:fig-past-40-years) shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession.
+Part of our data are from the NCHS(National Center for Health Statistics)[@nvss], in particular, we acquired the US birth rate data for the past 40 years, Figure \@ref(fig:fig-past-40-years) shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession.
 We also generated data from U.S.
 Birth Rate 1950-2022[@birthrate].
-We selected top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
+We selected the top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
 TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT
@@
 ## Data Characteristics
+The NCHS data source is very detailed, it contains a lot of demographic and geographic information.
+For example, for each birth, the data set records the race of the father and the mother, this can help us to analyze the birth rate among different race groups.
+It also contains the age of the mother and father, this can be really helpful if we want to examine the trend of teenager pregnancy.
 
@@
 
-Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with prediction for the future.
+Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with the prediction for the future.
 The births rates of China and South Africa were obviously higher than the U.S.
@@
 However, around the year of recession, the birth rate of China and Japan dropped even harder than the U.S.
-on the absolute level that are lower than 10 births per 1000 people.
+on the absolute level that is lower than 10 births per 1000 people.
 
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional details
-
-\newpage
-
 # References

### 2022-02-28 outputs/paper/paper.Rmd
fixed conflict, add abstract

@@
 Birth Rate 1950-2022[@birthrate].
-We selected the top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
+We selected the top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads.
+These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
 TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT

### 2022-02-28 outputs/paper/paper.Rmd
Update LGBT analysis

@@
 ---
-title: "What Happened to The Birth Rate?"
-subtitle: "The Great Recession and the LGBT Movement"
+title: "My title"
+subtitle: "My subtitle if needed"
 author: 
@@
 abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\".
-^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
+^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
 However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years.
@@
 
-The Great Recession has been a worldwide crisis that reconstructed the economy and was deep, synchronous and global.
-Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent, and that of Japan by more than 8 percent.
+The Great Recession has been a world-wide crisis that reconstructed the economy, and was deep, synchronous and global.
+Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent and that of Japan by more than 8 percent.
 The impact on international trade has been severe that all G20 countries experienced serious reductions in trade[@recession].
-Apart from the economic impact, the great recession also had a negative impact on the country's birth rate.
+Apart from economic impact, the great recession also had a negative impact on country's birth rate.
 
 In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas.
-Another impact from the depressed economic condition is delayed childbearing, and others to hasten it.
+Another impact from depressed economic condition is delayed childbearing, and others to hasten it.
 The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility].
@@
 
-As mentioned by @cite_original, the basic approach to many economic problems is maximizing utility.
-This approach recognized that children bring people utility but also comes with associated costs including time and money.
+As mentioned by @cite_original, the basic approach to many economic problem is maximizing utility.
+This approach recognized that children bring people utility but also with associated costs including time and money.
 This approach to modeling the decision to have a child leads to standard predictions of price and income effects.
 Direct consideration of costs could be like housing and medical care.
-Other times the price effect is indirect, as the opportunity cost of a woman’s time.
+Other times the price effect is indirect, like the opportunity cost of a woman’s time.
 Holding prices and quality constant, an increase in income will lead people to choose to have more children.
 This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per-capita GDP observed over time and place[@cite_original].
-However, this is enough for us to consider whether the birth rates are worldwide affected in the same negative direction after the Great Recession.
+However, this is enough for us to consider whether the birth rates are world-wide affected in the same negative direction after the Great Recession.
 
-Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started in the 1920s and had huge progress after the 2000s.
-A lot of laws that prohibited homosexual relationships were lifted and homosexual couples can get married legally and adopt kids, there is also propaganda that encourages homosexuality.
+Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started at the 1920s and had huge progress after 2000s.
+A lot of laws that prohibited homosexual relationship were lifted and homosexual couples can get married legally and adopt kids, there are also propaganda that encourages homosexuality.
 
-In this paper, we will investigate the effect of the Great Recession on the birth rate in different countries, and we will also explore the effect of the LGBT movement on the birth rate.
+In this paper, we will investigate the effect of the Great Recession on the birth rate in different countries, and we will also explore the effect of LGBT movement on birth rate.
 
@@
 ## Data Source
-Part of our data are from the NCHS(National Center for Health Statistics)[@nvss], in particular, we acquired the US birth rate data for the past 40 years, Figure \@ref(fig:fig-past-40-years) shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession.
+Part of our data are from the NCHS(National Center for Health Statistics)[@nvss], in particular, we acquired US birth rate data for the past 40 years, Figure \@ref(fig:fig-past-40-years) shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession.
 We also generated data from U.S.
 Birth Rate 1950-2022[@birthrate].
-We selected the top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads.
-These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
+We selected top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
 TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT
@@
 ## Data Characteristics
-The NCHS data source is very detailed, it contains a lot of demographic and geographic information.
-For example, for each birth, the data set records the race of the father and the mother, this can help us to analyze the birth rate among different race groups.
-It also contains the age of the mother and father, this can be really helpful if we want to examine the trend of teenager pregnancy.
 
@@
 
-Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with the prediction for the future.
+Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with prediction for the future.
 The births rates of China and South Africa were obviously higher than the U.S.
@@
 However, around the year of recession, the birth rate of China and Japan dropped even harder than the U.S.
-on the absolute level that is lower than 10 births per 1000 people.
+on the absolute level that are lower than 10 births per 1000 people.
 
@@
 
-## Effect of LGBT movement
+## Effect of LGBT Movement
 
@@
 
-Figure \@ref(fig:fig-gender-raising-children-distribution) shows the gender distribution of LGBT and Same Sex Couple as well as well as the percentage raising children.
-Figure \@ref(fig:fig-gender-raising-children-distribution) (a) and Figure \@ref(fig:fig-gender-raising-children-distribution) (c) clearly shows that more females are non-heterosexual than male are.
+Figure \@ref(fig:fig-gender-raising-children-distribution) shows the gender distribution of LGBT and Same Sex Couple as well as well as the percentage raising children [@will_institute].
+Figure \@ref(fig:fig-gender-raising-children-distribution) (a) and Figure \@ref(fig:fig-gender-raising-children-distribution) (c) clearly shows that more females are non-heterosexual than male are [@will_institute].
 This could cause a lower BirthRate-LGBT coefficient on birth rate as only females are able to give births (in normal situations).
@@
 
-Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US.
+Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US [@will_institute].
 Figure \@ref(fig:fig-density-state) [@cite_original] (c) shows the birth rate distribution by State in the US.
-Darker color means higher density or larger negative change in birth rate.
+Darker color means higher density or larger negative change in birth rate [@will_institute].
 From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate.
@@
 
-From \@ref(fig:fig-race-distribution) we can see that White population takes up the majority of the population of LGBT and Same Sex Couple.
-White is also the majority of the population of the US.
-Thus, larger LGBT population could lead to a even larger decrease in birth rate ($\phi<-1$, lower BirthRate-LGBT coefficient).
+From \@ref(fig:fig-race-distribution) we can see that White population takes up the majority of the population of LGBT (58%) and Same Sex Couple (74%) [@will_institute].
+White is also the majority of the population in the US [@quickfacts], around 76.3% of the total population.
+Thus, larger LGBT population could lead to an even larger decrease in birth rate ($\phi<-1$, lower BirthRate-LGBT coefficient).
 
@@
 
-From \@ref(fig:fig-age-distribution) (a) we can see that most same sex couples fall in the age group 25-64.
+From \@ref(fig:fig-age-distribution) (a) we can see that most same sex couples fall in the age group 25-64 [@will_institute].
 This period overlaps with the majority of the childbearing age.
-From \@ref(fig:fig-age-distribution) (b), we can see that LGBT population decreases as age increases starting from the age or 18, which is the beginning of women's childbearing age.
+From \@ref(fig:fig-age-distribution) (b) [@will_institute], we can see that LGBT population decreases as age increases starting from the age or 18, which is the beginning of women's childbearing age.
 Around 90% of the LGBT population have their age between 18 and 64, which covers almost all of the child bearing ages.
@@
 
-Figure \@ref(fig:fig-LGBT-socioeconomic-indicators) shows some Socioeconomic Indicators classified by sexual orientation (LGBT or not), such as employment status, income level education level, and whether or not being insured and food insecure.
+Figure \@ref(fig:fig-LGBT-socioeconomic-indicators) shows some socioeconomic indicators classified by sexual orientation (LGBT or not), such as employment status, income level education level, and whether or not being insured and food insecure [@will_institute].
 Darker color represents LGBT, lighter color represents non-LGBT.
 From the plots, we can extract some socioeconomic indicators of LGBT.
-Comparing to non-LGBT population, LGBT are more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is lower than that of the percentage of non-LGBT population).
+Comparing to non-LGBT population, LGBT are more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
+
+## Cohort Effect
 
@@
 
+## LGBT and Cohort Effect
+
 Figure \@ref(fig:fig-lgbt-trend-by-year) shows the trend of LGBT population percentage in total population.
-There is an obvious increasing trend (i.e. higher percentage of the US population are becoming LGBT). 
-Figure \@ref(fig:fig-lgbt-by-generation) shows the percentage of LGBT population trend in different generations.
+There is an obvious increasing trend (i.e. higher percentage of the US population are becoming LGBT) [@us_lgbt_trend].
+Figure \@ref(fig:fig-lgbt-by-generation) shows the percentage of LGBT population trend in different generations [@us_lgbt_trend].
 We can also see an increasing trend of LGBT population percentage (younger generations are more and more likely to become LGBT).
-This provide more evidence that cohort effects may contribute to the delines of birth rate in the US in recent years.
+This provides more evidence that cohort effects and the growth of LGBT population may have contributed a lot to the declines of birth rate in the US in recent years.
+
+\newpage
+
+\appendix
+
+# Appendix {-}
+
+# Additional details
 

### 2022-02-28 outputs/paper/paper.Rmd

--- 2022-02-28 outputs/paper/paper.Rmd
@@
 
+## LGBT and Cohort Effect
+
 Figure \@ref(fig:fig-lgbt-trend-by-year) shows the trend of LGBT population percentage in total population.
-There is an obvious increasing trend (i.e. higher percentage of the US population are becoming LGBT). 
-Figure \@ref(fig:fig-lgbt-by-generation) shows the percentage of LGBT population trend in different generations.
+There is an obvious increasing trend (i.e. higher percentage of the US population are becoming LGBT) [@us_lgbt_trend].
+Figure \@ref(fig:fig-lgbt-by-generation) shows the percentage of LGBT population trend in different generations [@us_lgbt_trend].
 We can also see an increasing trend of LGBT population percentage (younger generations are more and more likely to become LGBT).
-This provide more evidence that cohort effects may contribute to the delines of birth rate in the US in recent years.
+This provides more evidence that cohort effects and the growth of LGBT population may have contributed a lot to the declines of birth rate in the US in recent years.
 

### 2022-02-28 outputs/paper/paper.Rmd
pull

--- 2022-02-28 outputs/paper/paper.Rmd
@@
 
-We do observe declinations of world-wide birth rate despite of the impact level from recession.
-From the economic perspective, the countries who are least affected tend to have a stronger resistance to recession but not necessarily developed countries.
-While the most affected countries are mostly developing countries, they experienced a more significant plummet as percentage.
-It is also intriguing that the long-term births per 1000 people tend to converge to 10 for all countries mentioned above.
-However, the countries have higher resistance tend to delay the convergence which may derived from stronger economic status.
-Overall, the Great Recession was a turning point not only for the U.S.
-but also the whole world in the context of birth rate.
-The Great Recession may not be the direct cause of the world-wide declination in birth rates but it indeed provided an opportunity and provoked the desire for people to reconsider personal priorities with changed perspective of the society.
-
 ## Effect of LGBT movement
@@
 We can also see an increasing trend of LGBT population percentage (younger generations are more and more likely to become LGBT).
-This provide more evidence that cohort effects may contribute to the delines of birth rate in the US in recent years.
+This provide more evidence that cohort effects may contribute to the declines of birth rate in the US in recent years.
+
+# Discussion
+
+## Global Comparison
+
+We do observe declination of world-wide birth rate despite of the impact level from recession.
+From the economic perspective, the countries who are least affected tend to have a stronger resistance to recession but not necessarily developed countries.
+While the most affected countries are mostly developing countries, they experienced a more significant plummet as percentage.
+It is also intriguing that the long-term births per 1000 people tend to converge to 10 for all countries mentioned above.
+However, the countries have higher resistance tend to delay the convergence which may derived from stronger economic status.
+Overall, the Great Recession was a turning point not only for the U.S.
+but also the whole world in the context of birth rate.
+The Great Recession may not be the direct cause of the world-wide declination in birth rates but it indeed provided an opportunity and provoked the desire for people to reconsider personal priorities with changed perspective of the society.
 
 \newpage
-
 # References

### 2022-02-28 outputs/paper/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\title{What Happened to The Birth Rate?\thanks{Replication avaliable at \url{https://doi.org/10.48152/ssrp-mbxx-jm47}. Code and data avaliable at \url{https://github.com/cuilantao/Investigation-of-falling-US-Birth-Rates-since-the-great-recession}}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{The Great Recession and the LGBT Movement}
\author{Lantao Cui \and Huakun Shen \and Sijian Wu}
\date{28 February 2022}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={What Happened to The Birth Rate?},
  pdfauthor={Lantao Cui; Huakun Shen; Sijian Wu},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{subfig}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\begin{document}
\maketitle
\begin{abstract}
This paper is based on Kearney's paper ``The Puzzle of Falling US Birth Rates since the Great Recession.''\footnote{Original paper available at: \url{https://www.aeaweb.org/articles?id=10.1257/app.2.2.95}} (Kearney, Levine, and Pardue 2022) During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44. However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years. In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries. We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The Great Recession has been a worldwide crisis that reconstructed the economy and was deep, synchronous and global. Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent, and that of Japan by more than 8 percent. The impact on international trade has been severe that all G20 countries experienced serious reductions in trade(Michael Hart 2010). Apart from the economic impact, the great recession also had a negative impact on the country's birth rate.

In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas. Another impact from the depressed economic condition is delayed childbearing, and others to hasten it. The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases(Christine Percheski 2017). The significant consequence on fertility rate would make us wonder how severe is America being affected by the great recession. However, by comparing currency devaluations, equity market declines, and rising sovereign bond spreads, the U.S. is the third less affected country(URI DADUSH 2009).

Prior research suggests that stress may be linked to riskier sexual activity. Financial hardship and poverty can negatively affect cognitive function, and shorten the time horizons over which individuals make cost-benefit determinations. Alternatively, fertility may be unaffected by economic conditions if cultural norms related to the context and timing of births are particularly influential. For example, non-marital births are more socially consequential for some racial and ethnic groups, and in some geographic areas. For some individuals, these social norms may outweigh any economic considerations(Christine Percheski 2017). The same consideration of religious belief could also be applied to other countries.

As mentioned by Kearney, Levine, and Pardue (2022), the basic approach to many economic problems is maximizing utility. This approach recognized that children bring people utility but also comes with associated costs including time and money. This approach to modeling the decision to have a child leads to standard predictions of price and income effects. Direct consideration of costs could be like housing and medical care. Other times the price effect is indirect, as the opportunity cost of a woman's time. Holding prices and quality constant, an increase in income will lead people to choose to have more children. This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per-capita GDP observed over time and place(Kearney, Levine, and Pardue 2022). However, this is enough for us to consider whether the birth rates are worldwide affected in the same negative direction after the Great Recession.

Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started in the 1920s and had huge progress after the 2000s. A lot of laws that prohibited homosexual relationships were lifted and homosexual couples can get married legally and adopt kids, there is also propaganda that encourages homosexuality.

In this paper, we will investigate the effect of the Great Recession on the birth rate in different countries, and we will also explore the effect of the LGBT movement on the birth rate.

\hypertarget{data}{%
\section{Data}\label{data}}

We started our analysis by using R (R Core Team 2020), dplyr(Wickham et al. 2021), tidyverse(Wickham et al. 2019). Graphs are generated using ggpubr(Kassambara 2020) and ggplot2(Wickham 2016).

\hypertarget{data-source}{%
\subsection{Data Source}\label{data-source}}

Part of our data are from the NCHS(National Center for Health Statistics)({``NVSS - Birth Data''} 2022), in particular, we acquired the US birth rate data for the past 40 years, Figure \ref{fig:fig-past-40-years} shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession. We also generated data from U.S. Birth Rate 1950-2022(Trends 2022). We selected the top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads. These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth(URI DADUSH 2009). TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT

\hypertarget{data-characteristics}{%
\subsection{Data Characteristics}\label{data-characteristics}}

The NCHS data source is very detailed, it contains a lot of demographic and geographic information. For example, for each birth, the data set records the race of the father and the mother, this can help us to analyze the birth rate among different race groups. It also contains the age of the mother and father, this can be really helpful if we want to examine the trend of teenager pregnancy.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/fig-past-40-years-1.pdf}
\caption{\label{fig:fig-past-40-years}US Birth Rate in the past 40 years}
\end{figure}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{global-comparison-of-fertility-rate-under-the-great-recession}{%
\subsection{Global Comparison of Fertility Rate Under The Great Recession}\label{global-comparison-of-fertility-rate-under-the-great-recession}}

It shows that although the United States is at the epicenter of the global economic crisis, it is one of the countries least affected by the financial fallout. Large industrialized nations like the United States, Japan, and Germany have benefited from increasing global demand for relatively stable economies in which to invest. Instead, it is several developing countries, notably those with vulnerable capital accounts and weak macroeconomic fundamentals, that are experiencing severe economic downturns disproportionate to their roles in the crisis(URI DADUSH 2009).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/least-affected-1.pdf}
\caption{\label{fig:least-affected}Births per 1000 People of Top Least Affected Countries}
\end{figure}

Figure \ref{fig:least-affected} shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with the prediction for the future. The births rates of China and South Africa were obviously higher than the U.S. and Japan. However, around the year of recession, the birth rate of China and Japan dropped even harder than the U.S. on the absolute level that is lower than 10 births per 1000 people.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/most-affected-1.pdf}
\caption{\label{fig:most-affected}Births per 1000 People of Top Most Affected Countries}
\end{figure}

Figure \ref{fig:most-affected} shows the changes of births per 1000 people of Ukraine, Argentina, Hungary, Poland since 1950 with prediction for the future. All these most affected countries have lower initial births rate than those least affected countries. However, the levels of births per 1000 people were similar to those least affected countries around recession except Argentina. Ukraine, Hungary, and Poland also showed a potential increase after recession but continued to decline afterwards.

\textless\textless\textless\textless\textless\textless\textless{} HEAD
\#\# Effect of LGBT movement
=======
We do observe declinations of world-wide birth rate despite of the impact level from recession. From the economic perspective, the countries who are least affected tend to have a stronger resistance to recession but not necessarily developed countries. While the most affected countries are mostly developing countries, they experienced a more significant plummet as percentage. It is also intriguing that the long-term births per 1000 people tend to converge to 10 for all countries mentioned above. However, the countries have higher resistance tend to delay the convergence which may derived from stronger economic status. Overall, the Great Recession was a turning point not only for the U.S. but also the whole world in the context of birth rate. The Great Recession may not be the direct cause of the world-wide declination in birth rates but it indeed provided an opportunity and provoked the desire for people to reconsider personal priorities with changed perspective of the society.

\hypertarget{effect-of-lgbt-movement}{%
\subsection{Effect of LGBT Movement}\label{effect-of-lgbt-movement}}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
dce133630df142da93e120527ceec753178efac5
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

The percentage of LGBT (and many other gender identities) population has risen in recent years in the US, probably related to propaganda and political correctness.

Since there is not much data for other genders, and LGBT are the majorities, we will focus on LGBT population and the potential effect they bring to birth rate.

LGBT is the acronym of the following 4 gender identities, \textbf{L} for Lesbian, \textbf{G} for gay, \textbf{B} for bisexual, \textbf{T} for transgender. By definition, these genders should have a much lower birth rate than that of the heterosexual as normal births of human must depend on a male and a female.
Figure \ref{fig:fig-gender-raising-children-distribution} (b) shows the percentage of LGBT population that raise children. 71\% of the LGBT population don't raise children.
Then changes in birth rate and changes in LGBT population should have a negative proportional relationship (i.e.~Growth in LGBT population leads to lower birth rate).
Thus, the population changes in these population are very likely to make an impact on the birth rate of the US.

If we assume the simpliest scenario where we ignore race, gender distribution (equal percentage of male and female) and all other factors, and assume that no LGBT or same-sex-couple population give birth to a child, then there would be a perfect correlation (perfectly proportional) between LGBT population change and birth rate (i.e.~an \(x\%\) increase in LGBT population could lead to an \(x\%\) decrease in birth rate). Let's define \(\phi = \frac{\Delta Birth Rate}{\Delta LGBTPopulation}\) and call \(\phi\) the BirthRate-LGBT coefficient for convenience. Basically, lower BirthRate-LGBT coefficient means lower birth rate while holding LGBT population constant. If we take other factors into account, the impact to birth rate could be even larger (i.e.~\(\phi<-1\)).

\begin{figure}
\subfloat[LGBT Gender Distribution\label{fig:fig-gender-raising-children-distribution-1}]{\includegraphics[width=.33\linewidth]{./images/LGBT-Gender-distribution} }\subfloat[LGBT Percentage Raising Children\label{fig:fig-gender-raising-children-distribution-2}]{\includegraphics[width=.33\linewidth]{./images/LGBT-percentage-raising-children} }\subfloat[Same Sex Couple Gender Distribution\label{fig:fig-gender-raising-children-distribution-3}]{\includegraphics[width=.33\linewidth]{./images/same-sex-couple-gender-distribution} }\caption{Gender Distribution and Percentage Raising Children Distribution}\label{fig:fig-gender-raising-children-distribution}
\end{figure}

Figure \ref{fig:fig-gender-raising-children-distribution} shows the gender distribution of LGBT and Same Sex Couple as well as well as the percentage raising children (n.d.).
Figure \ref{fig:fig-gender-raising-children-distribution} (a) and Figure \ref{fig:fig-gender-raising-children-distribution} (c) clearly shows that more females are non-heterosexual than male are (n.d.).
This could cause a lower BirthRate-LGBT coefficient on birth rate as only females are able to give births (in normal situations).

\begin{figure}
\subfloat[Same Sex Couple State Density\label{fig:fig-density-state-1}]{\includegraphics[width=.49\linewidth]{./images/same-sex-couple-state-density} }\subfloat[LGBT State Density\label{fig:fig-density-state-2}]{\includegraphics[width=.49\linewidth]{./images/LGBT-state-density} }\newline\subfloat[Birth Rate Distribution by State\label{fig:fig-density-state-3}]{\includegraphics[width=.49\linewidth]{./images/state-birth-rate-distribution} }\caption{Density by State}\label{fig:fig-density-state}
\end{figure}

Figure \ref{fig:fig-density-state} (a) and Figure \ref{fig:fig-density-state} (b) shows the population density of Same Sex Couple and LGBT by States in the US (n.d.).
Figure \ref{fig:fig-density-state} (Kearney, Levine, and Pardue 2022) (c) shows the birth rate distribution by State in the US. Darker color means higher density or larger negative change in birth rate (n.d.). From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate. States on the west generally have darker color in all three plots. We cannot conclude whether LGBT population has a large or direct impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.

\begin{figure}

{\centering \subfloat[Same Sex Couple Race Distribution\label{fig:fig-race-distribution-1}]{\includegraphics[width=.30\linewidth]{./images/same-sex-couple-race-distribution} }\subfloat[LGBT Race Distribution\label{fig:fig-race-distribution-2}]{\includegraphics[width=.30\linewidth]{./images/LGBT-race-distribution} }

}

\caption{Race Distribution}\label{fig:fig-race-distribution}
\end{figure}

From \ref{fig:fig-race-distribution} we can see that White population takes up the majority of the population of LGBT (58\%) and Same Sex Couple (74\%) (n.d.). White is also the majority of the population in the US ({``U.s. Census Bureau Quickfacts: United States,''} n.d.), around 76.3\% of the total population. Thus, larger LGBT population could lead to an even larger decrease in birth rate (\(\phi<-1\), lower BirthRate-LGBT coefficient).

\begin{figure}

{\centering \subfloat[Same Sex Couple Race Distribution\label{fig:fig-age-distribution-1}]{\includegraphics[width=.40\linewidth]{./images/same-sex-couple-age-distribution} }\subfloat[LGBT Race Distribution\label{fig:fig-age-distribution-2}]{\includegraphics[width=.40\linewidth]{./images/LGBT-age-distribution} }

}

\caption{LGBT and Same Sex Couple Age Distribution}\label{fig:fig-age-distribution}
\end{figure}

From \ref{fig:fig-age-distribution} (a) we can see that most same sex couples fall in the age group 25-64 (n.d.). This period overlaps with the majority of the childbearing age. From \ref{fig:fig-age-distribution} (b) (n.d.), we can see that LGBT population decreases as age increases starting from the age or 18, which is the beginning of women's childbearing age. Around 90\% of the LGBT population have their age between 18 and 64, which covers almost all of the child bearing ages. These 2 evidences also indicate that, growth of LGBT population could lead to a larger decrease in birth rate (\(\phi < -1\), lower BirthRate-LGBT coefficient).

With these information, we can conclude that the growth of LGBT population is able to cause a negative change in birth rate.

\begin{figure}

{\centering \includegraphics[width=.80\linewidth]{./images/LGBT-socioeconomic-indicators}

}

\caption{Socioeconomic Indicators}\label{fig:fig-LGBT-socioeconomic-indicators}
\end{figure}

Figure \ref{fig:fig-LGBT-socioeconomic-indicators} shows some socioeconomic indicators classified by sexual orientation (LGBT or not), such as employment status, income level education level, and whether or not being insured and food insecure (n.d.). Darker color represents LGBT, lighter color represents non-LGBT.
From the plots, we can extract some socioeconomic indicators of LGBT. Comparing to non-LGBT population, LGBT are more likely to be unemployed, uninsured, being food insecure, having a low income (lower then \$24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).

\hypertarget{cohort-effect}{%
\subsection{Cohort Effect}\label{cohort-effect}}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/cohort-plot-1.pdf}
\caption{\label{fig:cohort-plot}Children Ever Born by Mother Age and Birth Cohort}
\end{figure}

Research has shown that cohort effects may have contributed to the declines of birth rate in the US(Kearney, Levine, and Pardue 2022).
``Period effects reflect changes that affect everyone at a point in time, whereas cohort effects reflect changes across people born or raised in different years.''(Kearney, Levine, and Pardue 2022).

We reproduced the plot showing Fertility Rate vs Mother's age from (Kearney, Levine, and Pardue 2022) in Figure \ref{fig:cohort-plot}. This plot is basically showing the fertility rate of women in different ages across 6 generations.
It's quite obvious that the 3 oldest cohorts (generations) have similar and higher fertility rates comparing to that of the 3 younger cohorts (generations).The 3 younger cohorts also have lower and lower fertility rate as their birthday moves forward (younger generation -\textgreater{} lower fertility rate).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/fig-parity-1.pdf}
\caption{\label{fig:fig-parity}Parity (ages 15-44)}
\end{figure}

Figure \ref{fig:fig-parity} is a replication of a plot showing trends in births by ``parity'' (referring to the number of children for a given woman) from (Kearney, Levine, and Pardue 2022). First and second births declined a lot since 2007, but the trend lines for third and higher order births are comparatively flat.
The women giving their 3rd and 4th births are most likely an older cohort than the cohorts giving their 1st and 2nd births. Again, this is another evidence showing the impact brought by cohort effects.

\begin{figure}

{\centering \includegraphics[width=.80\linewidth]{./images/US-LGBT-percent-trend}

}

\caption{Socioeconomic Indicators}\label{fig:fig-lgbt-trend-by-year}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=.80\linewidth]{./images/LGBT-percentage-by-generation}

}

\caption{Socioeconomic Indicators}\label{fig:fig-lgbt-by-generation}
\end{figure}

\hypertarget{lgbt-and-cohort-effect}{%
\subsection{LGBT and Cohort Effect}\label{lgbt-and-cohort-effect}}

Figure \ref{fig:fig-lgbt-trend-by-year} shows the trend of LGBT population percentage in total population.
There is an obvious increasing trend (i.e.~higher percentage of the US population are becoming LGBT) (Jones 2021).
Figure \ref{fig:fig-lgbt-by-generation} shows the percentage of LGBT population trend in different generations (Jones 2021).
We can also see an increasing trend of LGBT population percentage (younger generations are more and more likely to become LGBT).
This provides more evidence that cohort effects and the growth of LGBT population may have contributed a lot to the declines of birth rate in the US in recent years.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{global-comparison}{%
\subsection{Global Comparison}\label{global-comparison}}

We do observe declination of world-wide birth rate despite of the impact level from recession. From the economic perspective, the countries who are least affected tend to have a stronger resistance to recession but not necessarily developed countries. While the most affected countries are mostly developing countries, they experienced a more significant plummet as percentage. It is also intriguing that the long-term births per 1000 people tend to converge to 10 for all countries mentioned above. However, the countries have higher resistance tend to delay the convergence which may derived from stronger economic status. Overall, the Great Recession was a turning point not only for the U.S. but also the whole world in the context of birth rate. The Great Recession may not be the direct cause of the world-wide declination in birth rates but it indeed provided an opportunity and provoked the desire for people to reconsider personal priorities with changed perspective of the society.

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-will_institute}{}}%
n.d. \emph{The Williams Institute}. \url{https://williamsinstitute.law.ucla.edu/visualization/lgbt-stats/\#density}.

\leavevmode\vadjust pre{\hypertarget{ref-fertility}{}}%
Christine Percheski, Rachel Kimbro. 2017. {``How Did the Great Recession Affect Fertility?''} \url{https://www.irp.wisc.edu/publications/focus/pdfs/foc302g.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-us_lgbt_trend}{}}%
Jones, Jeffrey M. 2021. {``LGBT Identification Rises to 5.6.''} \emph{Gallup.com}. Gallup. \url{https://news.gallup.com/poll/329708/lgbt-identification-rises-latest-estimate.aspx}.

\leavevmode\vadjust pre{\hypertarget{ref-ggpubr}{}}%
Kassambara, Alboukadel. 2020. \emph{Ggpubr: 'Ggplot2' Based Publication Ready Plots}. \url{https://CRAN.R-project.org/package=ggpubr}.

\leavevmode\vadjust pre{\hypertarget{ref-cite_original}{}}%
Kearney, Melissa S., Phillip B. Levine, and Luke Pardue. 2022. {``The Puzzle of Falling US Birth Rates Since the Great Recession.''} \emph{Journal of Economic Perspectives} 36 (1): 151--76. \url{https://doi.org/10.1257/jep.36.1.151}.

\leavevmode\vadjust pre{\hypertarget{ref-recession}{}}%
Michael Hart, Bill Dymond. 2010. {``The Great Recession and International Trade.''} \url{https://policyoptions.irpp.org/magazines/g8g20/the-great-recession-and-international-trade/}.

\leavevmode\vadjust pre{\hypertarget{ref-nvss}{}}%
{``NVSS - Birth Data.''} 2022. \emph{Centers for Disease Control and Prevention}. Centers for Disease Control; Prevention. \url{https://www.cdc.gov/nchs/nvss/births.htm}.

\leavevmode\vadjust pre{\hypertarget{ref-r}{}}%
R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-birthrate}{}}%
Trends, Macro. 2022. {``Birth Rate 1950-2022.''} \url{https://www.macrotrends.net/countries/USA/united-states/birth-rate}.

\leavevmode\vadjust pre{\hypertarget{ref-rank}{}}%
URI DADUSH, SHIMELSE ALI, LAUREN FALCAO. 2009. {``The Unequal Impact of the Economic Crisis.''} \url{https://carnegieendowment.org/2009/07/09/unequal-impact-of-economic-crisis-pub-23385}.

\leavevmode\vadjust pre{\hypertarget{ref-quickfacts}{}}%
{``U.s. Census Bureau Quickfacts: United States.''} n.d. \emph{QuickFacts}. United States Census Bureau. \url{https://www.census.gov/quickfacts/fact/table/US/PST045221}.

\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

\leavevmode\vadjust pre{\hypertarget{ref-tidyverse}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. {``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2021. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.

\end{CSLReferences}

\end{document}

### 2022-02-28 outputs/paper/paper.Rmd
new change
Auto stash before merge of "main" and "origin/main"

--- 2022-02-28 outputs/paper/paper.Rmd
@@
 
-We do observe declinations of world-wide birth rate despite of the impact level from recession.
-From the economic perspective, the countries who are least affected tend to have a stronger resistance to recession but not necessarily developed countries.
-While the most affected countries are mostly developing countries, they experienced a more significant plummet as percentage.
-It is also intriguing that the long-term births per 1000 people tend to converge to 10 for all countries mentioned above.
-However, the countries have higher resistance tend to delay the convergence which may derived from stronger economic status.
-Overall, the Great Recession was a turning point not only for the U.S.
-but also the whole world in the context of birth rate.
-The Great Recession may not be the direct cause of the world-wide declination in birth rates but it indeed provided an opportunity and provoked the desire for people to reconsider personal priorities with changed perspective of the society.
-
 ## Effect of LGBT Movement
@@
 From the plots, we can extract some socioeconomic indicators of LGBT.
-Comparing to non-LGBT population, LGBT are more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
+Comparing to non-LGBT population, LGBT population might be more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
+It's not rigorous to conclude that LGBT are more
 
@@
 
+# Discussion
+
+## Global Comparison
+
+We do observe declination of world-wide birth rate despite of the impact level from recession.
+From the economic perspective, the countries who are least affected tend to have a stronger resistance to recession but not necessarily developed countries.
+While the most affected countries are mostly developing countries, they experienced a more significant plummet as percentage.
+It is also intriguing that the long-term births per 1000 people tend to converge to 10 for all countries mentioned above.
+However, the countries have higher resistance tend to delay the convergence which may derived from stronger economic status.
+Overall, the Great Recession was a turning point not only for the U.S.
+but also the whole world in the context of birth rate.
+The Great Recession may not be the direct cause of the world-wide declination in birth rates but it indeed provided an opportunity and provoked the desire for people to reconsider personal priorities with changed perspective of the society.
+
 \newpage
-
 # References

### 2022-02-28 outputs/paper/paper.tex
Auto stash before merge of "main" and "origin/main"

@@
 \author{Lantao Cui \and Huakun Shen \and Sijian Wu}
-\date{28 February 2022}
+\date{27 February 2022}
 
@@
 From the plots, we can extract some socioeconomic indicators of LGBT.
-Comparing to non-LGBT population, LGBT are more likely to be unemployed, uninsured, being food insecure, having a low income (lower then \$24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
+Comparing to non-LGBT population, LGBT population might be more likely to be unemployed, uninsured, being food insecure, having a low income (lower then \$24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
+It's not rigorous to conclude that LGBT are more
 

### 2022-02-28 outputs/paper/paper.Rmd
Auto stash before merge of "main" and "origin/main"

@@
   - \usepackage{subfig}
+  - \usepackage{float}
+  - \floatplacement{figure}{H}
 ---
@@
 Instead, it is several developing countries, notably those with vulnerable capital accounts and weak macroeconomic fundamentals, that are experiencing severe economic downturns disproportionate to their roles in the crisis[@rank].
-
+\newpage
 ```{r, least-affected, echo=FALSE, fig.cap="Births per 1000 People of Top Least Affected Countries"}
@@
 Comparing to non-LGBT population, LGBT population might be more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
-It's not rigorous to conclude that LGBT are more
+It's not rigorous to conclude for certain that LGBT are more likely to have these characteristics without a more in-depth statistical analysis, as there may be statistically insignificant errors in data.
+However, as all characteristics related to low-income discussed here are related to LGBT, there must be some correlation between these socioeconomic indicators and LGBT population and it's also worth noticing if government plans to affect birth rate.
 
@@
 This plot is basically showing the fertility rate of women in different ages across 6 generations.
-It's quite obvious that the 3 oldest cohorts (generations) have similar and higher fertility rates comparing to that of the 3 younger cohorts (generations).The 3 younger cohorts also have lower and lower fertility rate as their birthday moves forward (younger generation -> lower fertility rate).
+It's quite obvious that the 3 oldest cohorts (generations) have similar and higher fertility rates comparing to that of the 3 younger cohorts (generations).The 3 younger cohorts also have lower and lower fertility rate as their birthday moves forward (younger generation $\Rightarrow$ lower fertility rate).
 

### 2022-02-28 outputs/paper/paper.Rmd
figure

@@
 Instead, it is several developing countries, notably those with vulnerable capital accounts and weak macroeconomic fundamentals, that are experiencing severe economic downturns disproportionate to their roles in the crisis[@rank].
+
 \newpage
-```{r, least-affected, echo=FALSE, fig.cap="Births per 1000 People of Top Least Affected Countries"}
+```{r least-affected, echo=FALSE, fig.cap="Births per 1000 People of Top Least Affected Countries"}
 #Generating time series graphs of births per 1000 people for each least affected country

### 2022-02-28 outputs/paper/paper.Rmd
Add lawmaking factors to LGBT analysis

@@
 
-## Effect of LGBT Movement
-
-The percentage of LGBT (and many other gender identities) population has risen in recent years in the US, probably related to propaganda and political correctness.
-
-Since there is not much data for other genders, and LGBT are the majorities, we will focus on LGBT population and the potential effect they bring to birth rate.
-
-LGBT is the acronym of the following 4 gender identities, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender.
-By definition, these genders should have a much lower birth rate than that of the heterosexual as normal births of human must depend on a male and a female.
-Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children. 71% of the LGBT population don't raise children. 
-Then changes in birth rate and changes in LGBT population should have a negative proportional relationship (i.e. Growth in LGBT population leads to lower birth rate).
-Thus, the population changes in these population are very likely to make an impact on the birth rate of the US.
-
-If we assume the simpliest scenario where we ignore race, gender distribution (equal percentage of male and female) and all other factors, and assume that no LGBT or same-sex-couple population give birth to a child, then there would be a perfect correlation (perfectly proportional) between LGBT population change and birth rate (i.e.
-an $x\%$ increase in LGBT population could lead to an $x\%$ decrease in birth rate).
-Let's define $\phi = \frac{\Delta Birth Rate}{\Delta LGBTPopulation}$ and call $\phi$ the BirthRate-LGBT coefficient for convenience.
-Basically, lower BirthRate-LGBT coefficient means lower birth rate while holding LGBT population constant.
-If we take other factors into account, the impact to birth rate could be even larger (i.e.
-$\phi<-1$).
-
-```{r echo = FALSE, fig-gender-raising-children-distribution, fig.cap='Gender Distribution and Percentage Raising Children Distribution', fig.subcap=c('LGBT Gender Distribution', 'LGBT Percentage Raising Children', 'Same Sex Couple Gender Distribution'), out.width='.33\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c("./images/LGBT-Gender-distribution.png",
-  "./images/LGBT-percentage-raising-children.png",
-  "./images/same-sex-couple-gender-distribution.png"))
-```
-
-Figure \@ref(fig:fig-gender-raising-children-distribution) shows the gender distribution of LGBT and Same Sex Couple as well as well as the percentage raising children [@will_institute].
-Figure \@ref(fig:fig-gender-raising-children-distribution) (a) and Figure \@ref(fig:fig-gender-raising-children-distribution) (c) clearly shows that more females are non-heterosexual than male are [@will_institute].
-This could cause a lower BirthRate-LGBT coefficient on birth rate as only females are able to give births (in normal situations).
-
-```{r echo = FALSE, fig-density-state, fig.cap='Density by State', fig.subcap=c('Same Sex Couple State Density', 'LGBT State Density', 'Birth Rate Distribution by State'), out.width='.49\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics(c("./images/same-sex-couple-state-density.png", "./images/LGBT-state-density.png", "./images/state-birth-rate-distribution.png"))
-```
-
-Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US [@will_institute].
-Figure \@ref(fig:fig-density-state) [@cite_original] (c) shows the birth rate distribution by State in the US.
-Darker color means higher density or larger negative change in birth rate [@will_institute].
-From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate.
-States on the west generally have darker color in all three plots.
-We cannot conclude whether LGBT population has a large or direct impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.
-
-```{r echo = FALSE, fig-race-distribution, fig.cap='Race Distribution', fig.subcap=c('Same Sex Couple Race Distribution', 'LGBT Race Distribution'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2, fig.align="center"}
-knitr::include_graphics(c("./images/same-sex-couple-race-distribution.png", "./images/LGBT-race-distribution.png"))
-```
-
-From \@ref(fig:fig-race-distribution) we can see that White population takes up the majority of the population of LGBT (58%) and Same Sex Couple (74%) [@will_institute].
-White is also the majority of the population in the US [@quickfacts], around 76.3% of the total population.
-Thus, larger LGBT population could lead to an even larger decrease in birth rate ($\phi<-1$, lower BirthRate-LGBT coefficient).
-
-```{r echo = FALSE, fig-age-distribution, fig.cap='LGBT and Same Sex Couple Age Distribution', fig.subcap=c('Same Sex Couple Race Distribution', 'LGBT Race Distribution'), out.width='.40\\linewidth', fig.asp=1, fig.ncol = 2, fig.align="center"}
-knitr::include_graphics(c("./images/same-sex-couple-age-distribution.png", "./images/LGBT-age-distribution.png"))
-```
-
-From \@ref(fig:fig-age-distribution) (a) we can see that most same sex couples fall in the age group 25-64 [@will_institute].
-This period overlaps with the majority of the childbearing age.
-From \@ref(fig:fig-age-distribution) (b) [@will_institute], we can see that LGBT population decreases as age increases starting from the age or 18, which is the beginning of women's childbearing age.
-Around 90% of the LGBT population have their age between 18 and 64, which covers almost all of the child bearing ages.
-These 2 evidences also indicate that, growth of LGBT population could lead to a larger decrease in birth rate ($\phi < -1$, lower BirthRate-LGBT coefficient).
-
-With these information, we can conclude that the growth of LGBT population is able to cause a negative change in birth rate.
-
-```{r echo = FALSE, fig-LGBT-socioeconomic-indicators, fig.cap='Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
-knitr::include_graphics("./images/LGBT-socioeconomic-indicators.png")
-```
-
-Figure \@ref(fig:fig-LGBT-socioeconomic-indicators) shows some socioeconomic indicators classified by sexual orientation (LGBT or not), such as employment status, income level education level, and whether or not being insured and food insecure [@will_institute].
-Darker color represents LGBT, lighter color represents non-LGBT.
-From the plots, we can extract some socioeconomic indicators of LGBT.
-Comparing to non-LGBT population, LGBT population might be more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
-It's not rigorous to conclude for certain that LGBT are more likely to have these characteristics without a more in-depth statistical analysis, as there may be statistically insignificant errors in data.
-However, as all characteristics related to low-income discussed here are related to LGBT, there must be some correlation between these socioeconomic indicators and LGBT population and it's also worth noticing if government plans to affect birth rate.
-
 ## Cohort Effect
@@
 It's quite obvious that the 3 oldest cohorts (generations) have similar and higher fertility rates comparing to that of the 3 younger cohorts (generations).The 3 younger cohorts also have lower and lower fertility rate as their birthday moves forward (younger generation $\Rightarrow$ lower fertility rate).
+This reflects the impact to birth rate brought by cohort effects.
 
@@
 Figure \@ref(fig:fig-parity) is a replication of a plot showing trends in births by "parity" (referring to the number of children for a given woman) from [@cite_original].
-First and second births declined a lot since 2007, but the trend lines for third and higher order births are comparatively flat.
-The women giving their 3rd and 4th births are most likely an older cohort than the cohorts giving their 1st and 2nd births.
+First and second births declined a lot since 2007, but the trend lines for the third and higher order births are comparatively flat.
+The women giving their 3rd and 4th births are most likely older cohorts comparing to the cohorts giving their 1st and 2nd births.
 Again, this is another evidence showing the impact brought by cohort effects.
 
-```{r echo = FALSE, fig-lgbt-trend-by-year, fig.cap='Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
+"The explanation of cohort effects lies more in the way these cohorts were raised or in how they experienced their childhoods [@cite_original], than about a particular policy or cost factor post-2007"[@cite_original].
+[@cite_original] claims that "this type of investigation is much harder to under-take with standard data and econometric techniques" and provided a some speculations, such as parents' expectation, resources (e.g.
+time and money) spent on children, etc.
+
+We would like to raise another hypothesis related to effects to children from the society, laws and social medias, that's LGBT movement.
+
+## Effect of LGBT Movement
+
+The percentage of LGBT (and many other gender identities) population has risen in recent years in the US \@ref(fig:fig-lgbt-trend-by-year), probably related to propaganda in social medias, lawmaking and political correctness.
+This research doesn't try to investigate the causes of LGBT population growth, and since there is not much data for other genders, and LGBT are the majorities, we will focus on LGBT population and the potential effect they brought to birth rate.
+
+"On February 12, 2004, the first same-sex marriage in the United States happened in San Francisco, California."
+"On May 17, 2004, Massachusetts became the first state and sixth jurisdiction in the world to legalize same-sex marriage."
+"In 2008, California and Connecticut both legalized same-sex marriage, followed by Iowa, Vermont, and New Hampshire.
+Up until 2012, legalization came through state courts, the enactment of state legislation, or the result of the decisions of federal courts.
+On November 6, 2012, Maine, Maryland and Washington became the first states to legalize same-sex marriage through popular vote."
+"On June 26, 2015, in the landmark civil rights case Obergefell v.
+Hodges, the Supreme Court ruled that the fundamental right to marry is guaranteed to same-sex couples by both the Due Process Clause and the Equal Protection Clause of the Fourteenth Amendment to the United States Constitution."[@same_sex_marriage_states_2022]
+The first two dates mentioned above represents the beginning of the legalization of same-sex marriage in the US, and is quite close to 2007, when the significant drop of birth rate began.
+There may be a composition of effects causing the significant drop of birth rate in 2007, and the legalization of same-sex population could potentially be one of main causes.
+
+LGBT is the acronym of the following 4 gender identities, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender.
+By definition, these genders should have a much lower birth rate than that of the heterosexual as normal births of human must depend on a male and a female.
+Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children. 71% of the LGBT population don't raise children. 
+Then changes in birth rate and changes in LGBT population should have a negative proportional relationship (i.e. Growth in LGBT population leads to lower birth rate).
+Thus, the population changes in these population are very likely to make an impact on the birth rate of the US.
+
+If we assume the simpliest scenario where we ignore race, gender distribution (equal percentage of male and female) and all other factors, and assume that no LGBT or same-sex-couple population give birth to a child, then there would be a perfect correlation (perfectly proportional) between LGBT population change and birth rate (i.e.
+an $x\%$ increase in LGBT population could lead to an $x\%$ decrease in birth rate).
+Let's define $\phi = \frac{\Delta Birth Rate}{\Delta LGBTPopulation}$ and call $\phi$ the BirthRate-LGBT coefficient for convenience.
+Basically, lower BirthRate-LGBT coefficient means lower birth rate while holding LGBT population constant.
+If we take other factors into account, the impact to birth rate could be even larger (i.e.
+$\phi<-1$).
+
+```{r echo = FALSE, fig-gender-raising-children-distribution, fig.cap='Gender Distribution and Percentage Raising Children Distribution', fig.subcap=c('LGBT Gender Distribution', 'LGBT Percentage Raising Children', 'Same Sex Couple Gender Distribution'), out.width='.33\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c("./images/LGBT-Gender-distribution.png",
+  "./images/LGBT-percentage-raising-children.png",
+  "./images/same-sex-couple-gender-distribution.png"))
+```
+
+Figure \@ref(fig:fig-gender-raising-children-distribution) shows the gender distribution of LGBT and Same Sex Couple as well as well as the percentage raising children [@will_institute].
+Figure \@ref(fig:fig-gender-raising-children-distribution) (a) and Figure \@ref(fig:fig-gender-raising-children-distribution) (c) clearly shows that more females are non-heterosexual than male are [@will_institute].
+This could cause a lower BirthRate-LGBT coefficient on birth rate as only females are able to give births (in normal situations).
+
+```{r echo = FALSE, fig-density-state, fig.cap='Density by State', fig.subcap=c('Same Sex Couple State Density', 'LGBT State Density', 'Birth Rate Distribution by State'), out.width='.49\\linewidth', fig.asp=1, fig.ncol = 2}
+knitr::include_graphics(c("./images/same-sex-couple-state-density.png", "./images/LGBT-state-density.png", "./images/state-birth-rate-distribution.png"))
+```
+
+Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US [@will_institute].
+Figure \@ref(fig:fig-density-state) [@cite_original] (c) shows the birth rate distribution by State in the US.
+Darker color means higher density or larger negative change in birth rate [@will_institute].
+From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate.
+States on the west generally have darker color in all three plots.
+We cannot conclude whether LGBT population has a large or direct impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.
+
+```{r echo = FALSE, fig-race-distribution, fig.cap='Race Distribution', fig.subcap=c('Same Sex Couple Race Distribution', 'LGBT Race Distribution'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2, fig.align="center"}
+knitr::include_graphics(c("./images/same-sex-couple-race-distribution.png", "./images/LGBT-race-distribution.png"))
+```
+
+From \@ref(fig:fig-race-distribution) we can see that White population takes up the majority of the population of LGBT (58%) and Same Sex Couple (74%) [@will_institute].
+White is also the majority of the population in the US [@quickfacts], around 76.3% of the total population.
+Thus, larger LGBT population could lead to an even larger decrease in birth rate ($\phi<-1$, lower BirthRate-LGBT coefficient).
+
+```{r echo = FALSE, fig-age-distribution, fig.cap='LGBT and Same Sex Couple Age Distribution', fig.subcap=c('Same Sex Couple Race Distribution', 'LGBT Race Distribution'), out.width='.40\\linewidth', fig.asp=1, fig.ncol = 2, fig.align="center"}
+knitr::include_graphics(c("./images/same-sex-couple-age-distribution.png", "./images/LGBT-age-distribution.png"))
+```
+
+From \@ref(fig:fig-age-distribution) (a) we can see that most same sex couples fall in the age group 25-64 [@will_institute].
+This period overlaps with the majority of the childbearing age.
+From \@ref(fig:fig-age-distribution) (b) [@will_institute], we can see that LGBT population decreases as age increases starting from the age or 18, which is the beginning of women's childbearing age.
+Around 90% of the LGBT population have their age between 18 and 64, which covers almost all of the child bearing ages.
+These 2 evidences also indicate that, growth of LGBT population could lead to a larger decrease in birth rate ($\phi < -1$, lower BirthRate-LGBT coefficient).
+
+With these information, we can conclude that the growth of LGBT population is able to cause a negative change in birth rate.
+
+```{r echo = FALSE, fig-LGBT-socioeconomic-indicators, fig.cap='Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
+knitr::include_graphics("./images/LGBT-socioeconomic-indicators.png")
+```
+
+Figure \@ref(fig:fig-LGBT-socioeconomic-indicators) shows some socioeconomic indicators classified by sexual orientation (LGBT or not), such as employment status, income level education level, and whether or not being insured and food insecure [@will_institute].
+Darker color represents LGBT, lighter color represents non-LGBT.
+From the plots, we can extract some socioeconomic indicators of LGBT.
+Comparing to non-LGBT population, LGBT population might be more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
+It's not rigorous to conclude for certain that LGBT are more likely to have these characteristics without a more in-depth statistical analysis, as there may be statistically insignificant errors in data.
+However, as all characteristics related to low-income discussed here are related to LGBT, there must be some correlation between these socioeconomic indicators and LGBT population and it's also worth noticing if government plans to affect birth rate.
+
+```{r echo = FALSE, fig-lgbt-trend-by-year, fig.cap='LGBT Population Percentage By Year', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
 knitr::include_graphics("./images/US-LGBT-percent-trend.png")
@@
 
-```{r echo = FALSE, fig-lgbt-by-generation, fig.cap='Socioeconomic Indicators', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
+```{r echo = FALSE, fig-lgbt-by-generation, fig.cap='LGBT Population Percentage By Generation', fig.asp=1, fig.ncol = 1, out.width='.80\\linewidth', fig.align="center"}
 knitr::include_graphics("./images/LGBT-percentage-by-generation.png")

### 2022-02-28 outputs/paper/paper.Rmd
Add data source for LGBT related analysis

@@
 These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
-TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT
+
+The data used to analyze the impact to birth rate brought by LGBT population growth comes from School of Law, UCLA[@will_institute].
+The analysis done be School of Law, UCLA[@will_institute] were from the data collected for the Gallup Daily Tracking survey.
+
+The data used to analyze LGBT population trends comes from GALLUP News [@us_lgbt_trend], which is the source of data of the LGBT Demographics analysis.
+The data is based on telephone interviews conducted in 2020 with a random sample of 15,349 adults (18+), living in the 50 states and the District of Columbia.
+The entire sample of national adults has a margin of sampling error is $\pm1\%$ at the $95\%$ confidence level.
+The results based on the total sample of 692 LGBT adults has a sample error of $\pm5\%$ with a $95\%$ confidence level.
 
@@
 It also contains the age of the mother and father, this can be really helpful if we want to examine the trend of teenager pregnancy.
+
+The LGBT demographics data spans 2012 to 2017.
+The sample is a approximately 350,000 U.S.
+adults (18+) who resides in the 50 states and district of Columbia.
+The data describes LGBT Data & Demographics such as LGBT proportion of population by States, characteristics of LGBT people (race, gender distribution, etc.), socioeconomic indicators (income, education level, etc.) and State comparison.
+
+The GALLUP News data used to analyze LGBT population trend shows the LGBT population proportion and distribution trend from 2012 to 2020 by year and generation, as well as the distribution of sexual orientation of different gender identity.
 

### 2022-02-28 outputs/paper/paper.Rmd
Changed minor details, typos, wording

--- 2022-02-28 outputs/paper/paper.Rmd
@@
 abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\".
-^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44.
+^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 births per 1000 women between the age of 15 to 44.
 However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years.
 In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries.
-We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect"
+We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect."
 output:
@@
 Another impact from the depressed economic condition is delayed childbearing, and others to hasten it.
-The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility].
+The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.
+There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility].
 The significant consequence on fertility rate would make us wonder how severe is America being affected by the great recession.
@@
 
-Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started in the 1920s and had huge progress after the 2000s.
+Another factor other than policy or economic factors that might have affected the fertility rate is cohort effect, more specifically, in this paper we focus on the rise of the LGBT community.
+The LGBT activity started in the 1920s and had a huge progress after the 2000s[@same_sex_marriage_states_2022].
 A lot of laws that prohibited homosexual relationships were lifted and homosexual couples can get married legally and adopt kids, there is also propaganda that encourages homosexuality.
@@
 These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].
-TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT
+
+The data used to analyze the impact to birth rate brought by LGBT population growth comes from the School of Law, UCLA[@will_institute].
+The analysis was based on the data collected for the Gallup Daily Tracking survey.
+
+The data used to analyze LGBT population trends comes from GALLUP News [@us_lgbt_trend], which is the source of data of the LGBT Demographics analysis[@will_institute].
+The data is based on telephone interviews conducted in 2020 with a random sample of 15,349 adults (18+), living in the 50 states and the District of Columbia.
+The entire sample of national adults has a margin of sampling error is $\pm1\%$ at the $95\%$ confidence level.
+The results based on the total sample of 692 LGBT adults has a sample error of $\pm5\%$ with a $95\%$ confidence level.
 
@@
 It also contains the age of the mother and father, this can be really helpful if we want to examine the trend of teenager pregnancy.
+
+The LGBT demographics data spans 2012 to 2017.
+The sample is approximately 350,000 U.S.
+adults (18+) who resides in the 50 states and district of Columbia.
+The data describes LGBT Data & Demographics such as LGBT proportion of population by states, characteristics of LGBT people (race, gender distribution, etc.), socioeconomic indicators (income, education level, etc.) and state comparison.
+
+The GALLUP News data used to analyze LGBT population trend shows the LGBT population proportion and distribution trend from 2012 to 2020 by year and generation, as well as the distribution of sexual orientation of different gender identities.
 
@@
 Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with the prediction for the future.
-The births rates of China and South Africa were obviously higher than the U.S.
+The birth rates of China and South Africa were obviously higher than that of the U.S.
 and Japan.
@@
 Figure \@ref(fig:most-affected) shows the changes of births per 1000 people of Ukraine, Argentina, Hungary, Poland since 1950 with prediction for the future.
-All these most affected countries have lower initial births rate than those least affected countries.
+All these most affected countries have lower initial birth rates than that of the those least affected countries.
 However, the levels of births per 1000 people were similar to those least affected countries around recession except Argentina.
@@
 
-Figure \@ref(fig:fig-parity) is a replication of a plot showing trends in births by "parity" (referring to the number of children for a given woman) from [@cite_original].
-First and second births declined a lot since 2007, but the trend lines for the third and higher order births are comparatively flat.
+Figure \@ref(fig:fig-parity) is a reproduction of a plot showing trends in births by "parity" (referring to the number of children for a given woman) from [@cite_original].
+The first and second births declined a lot since 2007, but the trend lines for the third and higher order births are comparatively flat.
 The women giving their 3rd and 4th births are most likely older cohorts comparing to the cohorts giving their 1st and 2nd births.

### 2022-02-28 outputs/paper/paper.Rmd
backup progress

@@
 title: "What Happened to The Birth Rate?"
-subtitle: "The Great Recession and the LGBT Movement"
+subtitle: "Impact to Birth Rate Brought By The Great Recession and The LGBT Movement"
 author: 
@@
 However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years.
-In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries.
-We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect."
+Maintaining the fertility rate to be around 2.1 (replacement level) is important because low fertility rate could lead to aging of population and economic stagnation of economy.
+In this paper, we expand our research on multiple countries and found out that not only the US was affected by the recession, but the whole World in the context of birth rate, and the most affected countries are mostly developing countries.
+We also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect, and found that the growth of LGBT population may have a cohort effect and lead caused a negative impact on the birth rate."
 output:
@@
 
-"The explanation of cohort effects lies more in the way these cohorts were raised or in how they experienced their childhoods [@cite_original], than about a particular policy or cost factor post-2007"[@cite_original].
-[@cite_original] claims that "this type of investigation is much harder to under-take with standard data and econometric techniques" and provided a some speculations, such as parents' expectation, resources (e.g.
-time and money) spent on children, etc.
-
-We would like to raise another hypothesis related to effects to children from the society, laws and social medias, that's LGBT movement.
-
 ## Effect of LGBT Movement
 
-The percentage of LGBT (and many other gender identities) population has risen in recent years in the US \@ref(fig:fig-lgbt-trend-by-year), probably related to propaganda in social medias, lawmaking and political correctness.
-This research doesn't try to investigate the causes of LGBT population growth, and since there is not much data for other genders, and LGBT are the majorities, we will focus on LGBT population and the potential effect they brought to birth rate.
+### Laws for Same Sex Marriage
 
@@
 "On June 26, 2015, in the landmark civil rights case Obergefell v.
-Hodges, the Supreme Court ruled that the fundamental right to marry is guaranteed to same-sex couples by both the Due Process Clause and the Equal Protection Clause of the Fourteenth Amendment to the United States Constitution."[@same_sex_marriage_states_2022]
+Hodges, the Supreme Court ruled that the fundamental right to marry is guaranteed to same-sex couples by both the Due Process Clause and the Equal Protection Clause of the Fourteenth Amendment to the United States Constitution."[@same_sex_marriage_states_2022].
+
 The first two dates mentioned above represents the beginning of the legalization of same-sex marriage in the US, and is quite close to 2007, when the significant drop of birth rate began.
@@
 
+### LGBT Demographics
+
 LGBT is the acronym of the following 4 gender identities, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender.
 By definition, these genders should have a much lower birth rate than that of the heterosexual as normal births of human must depend on a male and a female.
-Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children. 71% of the LGBT population don't raise children. 
 Then changes in birth rate and changes in LGBT population should have a negative proportional relationship (i.e. Growth in LGBT population leads to lower birth rate).
@@
 
-```{r echo = FALSE, fig-density-state, fig.cap='Density by State', fig.subcap=c('Same Sex Couple State Density', 'LGBT State Density', 'Birth Rate Distribution by State'), out.width='.49\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics(c("./images/same-sex-couple-state-density.png", "./images/LGBT-state-density.png", "./images/state-birth-rate-distribution.png"))
-```
-
-Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US [@will_institute].
-Figure \@ref(fig:fig-density-state) [@cite_original] (c) shows the birth rate distribution by State in the US.
-Darker color means higher density or larger negative change in birth rate [@will_institute].
-From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate.
-States on the west generally have darker color in all three plots.
-We cannot conclude whether LGBT population has a large or direct impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.
-
 ```{r echo = FALSE, fig-race-distribution, fig.cap='Race Distribution', fig.subcap=c('Same Sex Couple Race Distribution', 'LGBT Race Distribution'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2, fig.align="center"}
@@
 
-With these information, we can conclude that the growth of LGBT population is able to cause a negative change in birth rate.
+### Socioeconomic Inicators
 
@@
 Comparing to non-LGBT population, LGBT population might be more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).
-It's not rigorous to conclude for certain that LGBT are more likely to have these characteristics without a more in-depth statistical analysis, as there may be statistically insignificant errors in data.
-However, as all characteristics related to low-income discussed here are related to LGBT, there must be some correlation between these socioeconomic indicators and LGBT population and it's also worth noticing if government plans to affect birth rate.
+
+### Geographical Distribution of LGBT Population and Birth Rate
+
+```{r echo = FALSE, fig-density-state, fig.cap='Density by State', fig.subcap=c('Same Sex Couple State Density', 'LGBT State Density', 'Birth Rate Distribution by State'), out.width='.49\\linewidth', fig.asp=1, fig.ncol = 2}
+knitr::include_graphics(c("./images/same-sex-couple-state-density.png", "./images/LGBT-state-density.png", "./images/state-birth-rate-distribution.png"))
+```
+
+Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US [@will_institute].
+Figure \@ref(fig:fig-density-state) (c) [@cite_original] shows the birth rate distribution by state in the US.
+Darker color means higher density or larger negative change in birth rate [@will_institute].
+From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate.
+States on the west generally have darker color in all three plots.
+We cannot conclude whether LGBT population has a large or direct impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.
 
@@
 ```
-
-## LGBT and Cohort Effect
-
-Figure \@ref(fig:fig-lgbt-trend-by-year) shows the trend of LGBT population percentage in total population.
-There is an obvious increasing trend (i.e. higher percentage of the US population are becoming LGBT) [@us_lgbt_trend].
-Figure \@ref(fig:fig-lgbt-by-generation) shows the percentage of LGBT population trend in different generations [@us_lgbt_trend].
-We can also see an increasing trend of LGBT population percentage (younger generations are more and more likely to become LGBT).
-This provides more evidence that cohort effects and the growth of LGBT population may have contributed a lot to the declines of birth rate in the US in recent years.
 
@@
 
+## Cohort Effects and LGBT Movement Effects
+
+With the results described in the **[Cohort Effect]** section, we can see that cohort effect can largely decrease the birth rate.
+
+"The explanation of cohort effects lies more in the way these cohorts were raised or in how they experienced their childhoods [@cite_original], than about a particular policy or cost factor post-2007"[@cite_original].
+[@cite_original] claims that "this type of investigation is much harder to under-take with standard data and econometric techniques" and provided a some speculations, such as parents' expectation, resources (e.g.
+time and money) spent on children, etc.
+
+We would like to raise another hypothesis related to effects to children from the society, laws and social medias, that's LGBT movement.
+
+The percentage of LGBT (and many other gender identities) population has risen in recent years in the US \@ref(fig:fig-lgbt-trend-by-year), probably related to propaganda in social medias, lawmaking and political correctness.
+This analysis doesn't try to investigate the causes of LGBT population growth, and since there is not much data for other genders, and LGBT are the majorities, we will focus on LGBT population and the potential effect they brought to birth rate.
+
+With the results described in the **[Effect of LGBT Movement]** section, we showed that many factors related to LGBT population growth is able to cause a negative change in birth rate.
+
+In section **[Socioeconomic Inicators]** we also discussed some socioeconomic characteristics of the LGBT population.
+The characteristics are all related to low-income or low-education level.
+However, it is not rigorous to conclude for certain that LGBT are more likely to have these characteristics without a more in-depth statistical analysis, as there may be statistically insignificant errors in data.
+However, as all characteristics related to low-income discussed here are related to LGBT, there must be some correlation between these socioeconomic indicators and LGBT population and it's also worth noticing if government wishes to adjust birth rate.
+
+Figure \@ref(fig:fig-lgbt-trend-by-year) shows the trend of LGBT population percentage in total population.
+There is an obvious increasing trend (i.e. higher percentage of the US population are becoming LGBT) [@us_lgbt_trend].
+Figure \@ref(fig:fig-lgbt-by-generation) shows the percentage of LGBT population trend in different generations [@us_lgbt_trend].
+We can also see an increasing trend of LGBT population percentage (younger generations are more and more likely to become LGBT).
+This provides more evidence that cohort effects and the growth of LGBT population may have contributed a lot to the declines of birth rate in the US in recent years.
+
 \newpage

### 2022-02-28 outputs/paper/paper.Rmd

@@
 Maintaining the fertility rate to be around 2.1 (replacement level) is important because low fertility rate could lead to aging of population and economic stagnation of economy.
-In this paper, we expand our research on multiple countries and found out that not only the US was affected by the recession, but the whole World in the context of birth rate, and the most affected countries are mostly developing countries.
-We also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect, and found that the growth of LGBT population may have a cohort effect and lead caused a negative impact on the birth rate."
+In this paper, we expanded our research on multiple countries and found out that not only the US was affected by the recession, but the whole World in the context of birth rate, and the most affected countries are mostly developing countries.
+We also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examined its effect, and found that the growth of LGBT population may have a cohort effect and lead caused a negative impact on the birth rate."
 output:
@@
 A lot of laws that prohibited homosexual relationships were lifted and homosexual couples can get married legally and adopt kids, there is also propaganda that encourages homosexuality.
+Birth rates and LGBT was found to be highly correlated, the growth of LGBT population could lead to lower birth rate and the growth of LGBT population is increasing rapidly in the younger generations.
 
@@
 The analysis was based on the data collected for the Gallup Daily Tracking survey.
+Respondents were obtained through list-assisted random digit dial (70% cell phone, 30% landline).
 
@@
 ```{r, fig-past-40-years, fig.cap="US Birth Rate in the past 40 years", echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=4}
+# import the csv
 US_birth_date_from_1985_to_2020 <- read_csv(here::here("inputs/data/Birthrate.csv"))
 
+# Plot the graph
 ggplot(data=US_birth_date_from_1985_to_2020, aes(x=`year`, y=`brate_all`, group=1)) +
@@
   xlab("Year") + ylab("Births per 1,000 women age 15-44")
-
 ```
@@
 ```{r cohort-plot, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.cap="Children Ever Born by Mother Age and Birth Cohort", fig.height=4, fig.width=8}
+# Read the dta
 age_comp_pop <- read_dta(here::here("inputs/data/agecomp-seer.dta"))
+# Filter out unwanted columns
 age_comp_pop <- subset(age_comp_pop, select = -c(fem1519, fem2034, fem3544, stname))
+# group by and find the sum of all columns
 age_comp_pop_grouped <- group_by(age_comp_pop, year) %>% summarise(across(everything(), sum))
+# generate vector of column names that start with fem
 age_comp_pop_grouped_start_with_fem <-
@@
 fem_col_name <- colnames(age_comp_pop_grouped_start_with_fem)
+# transform wide to long
 age_comp_pop_long <- gather(age_comp_pop_grouped, mage, fem, fem_col_name)
+# generate new columns
 age_comp_pop_long$mage <- gsub("^.{0,3}", "", age_comp_pop_long$mage)
+# transform it into a numeric value
 age_comp_pop_long$mage <- as.numeric(as.character(age_comp_pop_long$mage))
@@
 age_comp_pop_long$cohort2 <- ''
-
+# assign different numbers based on years
 age_comp_pop_long$cohort2[age_comp_pop_long$cohort >= 1968 & age_comp_pop_long$cohort <= 1972] <- 1
@@
 age_comp_pop_long_group <- group_by(age_comp_pop_long, cohort2, mage) %>% summarise(pop_sum = sum(pop))
+# read data files
 cohort_analysis <- read_dta(here::here("inputs/data/nchs_cohort_analysis.dta"))
+# merge two file together
 merged_data <- merge(cohort_analysis, age_comp_pop_long_group, by = c("cohort2", "mage"))
+# define new columns
 merged_data$brate <- with(merged_data, numbirth/pop_sum*1000)
@@
 data_with_only_cohort_and_brate <- data_with_only_cohort_and_brate[order(data_with_only_cohort_and_brate$cohort2, data_with_only_cohort_and_brate$mage),]
+# accmulate all the values
 data_with_only_cohort_and_brate$brate <- ave(data_with_only_cohort_and_brate$brate, data_with_only_cohort_and_brate$cohort2, FUN=cumsum)
@@
 final_data_wide_format <- spread(data_with_only_cohort_and_brate_grouped, key = cohort, value = cum_brate,  sep = '')
-
+# drop all na value
 data_with_cohort1 <- drop_na(final_data_wide_format, cohort1)
@@
 data_with_cohort6 <- drop_na(final_data_wide_format, cohort6)
-
+# draw the plot
 ggplot() +
@@
 ```{r fig-parity, fig.cap="Parity (ages 15-44)", echo=FALSE, fig.height=4, fig.height=3}
+# Import the dta
 birth_pop <- read_dta(here::here("inputs/data/nchs_births_pop_1990_2019.dta"))
 age_race_pop <- read_dta(here::here("inputs/data/age_race_comp_seer.dta"))
-
 merged_data <- merge(birth_pop, age_race_pop, by = c("stname", "year"))
@@
 merged_data$pop2044 <- with(merged_data, `pop2024.y` + `pop2534.y` + `pop3544.y`)
-
+# use a group by to get the sum
 merged_data_grouped_by <- group_by(merged_data, year) %>% summarise(fb_sum = sum(numbirth_firstbirth), sb_sum = sum(numbirth_secondbirth), tb_sum = sum(numbirth_thirdbirth), fbp_sum = sum(numbirth_fourthbirth), pop_sum = sum(pop1544))
-
+#define new columns based on existing columns
 merged_data_grouped_by$brith_rate_fb <- with(merged_data_grouped_by, fb_sum / `pop_sum` * 1000)
@@
 merged_data_grouped_by$brith_rate_fbp <- with(merged_data_grouped_by, fbp_sum / `pop_sum` * 1000)
-
+# merge the plot and add legends
 ggplot(merged_data_grouped_by, aes(x=year)) + 
